---
title: "Analysis of L&P dataset"
output:
  html_document: default
  word_document: default
---

```{r LoadLibs, echo=FALSE, include=FALSE}
library(xtable)
library(knitr)
library(RSQLite)
library(lubridate)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library (hydroGOF)
library(scales)
library(GGally)

options("scipen"=100, "digits"=4) # prints not in scientific notation

```

## FIXME list

- Print numeric value of Surfacekl in csv output
- Use numeric value of parameters as characters througout to simplify comparisons
- Depth can come as numeric too
- The current criteria for parameter selection has HUGE leverage on results
- How to define best fit?????

## Load and massage the data

- Tidy up factors/levels (note the need to set new label on csv dataset)

- Separete in Sedling and regrowth phases

- Get data formats right


```{r LoadTidyUp, echo=FALSE,include=FALSE, echo=FALSE}

#dataRaw <- read.csv("C:\\GitHubRepos\\LP_analysis\\ObsPre.csv", header = TRUE, skip = 4)
dataRaw <- read.csv("K:\\CPDiary\\Data\\Lucerne & pasture drought trial 2011-2012\\ModelFitting\\ObsPre.csv", 
                    header = TRUE)
head(dataRaw)
summary(dataRaw)

# Create working df and sort out formats - 
dataWork <- as.data.frame(dataRaw)

# Force parameter values to factors
dataWork$Date <- as.Date(ymd(dataWork$Date))
dataWork$SurfaceKl <- as.factor(dataWork$SurfaceKl)
dataWork$RootFrontVelocity <- as.factor(dataWork$RootFrontVelocity)
dataWork$KLReduction <- as.factor(dataWork$KLReduction)

# sort depth out
dataWork$Depth <- gsub("D", "", dataWork$Depth)
dataWork$Depth <- factor(dataWork$Depth, 
                         levels = c(7, 23, 30, 50, 70, 90, 110, 130, 150))
dataWork$Depth <- as.factor(dataWork$Depth)

# Create separate Establishment and Regrowth periods
estDate <- "2012-06-01" # FIXME: find a more functional criteria later
#estDate <- "2012-01-01" # FIXME: test sensitivity to this date
#estDate <- "2012-06-01 00:00:00"
dataWork$Stage <- as.factor(ifelse(dataWork$Date<=estDate,"Establishment","Regrowth"))
dataWork <- dataWork %>% mutate(Irrigation = factor(Irrigation, 
                                                    levels=c("None", 
                                                             "ThreeWeekly",
                                                             "TwoPerWeek",
                                                             "OnePerWeek"))) # re-order factors

levels(dataWork$Irrigation) <- c("No Irrigation",
                                 "Low Frequency",
                                 "Medium Frequency", 
                                 "High Frequency") # rename factors

# Check levels
levels(dataWork$Depth)
levels(dataWork$SurfaceKl)
levels(dataWork$KLReduction)
levels(dataWork$RootFrontVelocity)

# Check
head(dataWork)
summary(dataWork)
str(dataWork)

```


## Check clean dataset


```{r, echo=FALSE}
# final product
str(dataWork)

```

```{r, include = FALSE}
head(dataWork, 20)
```

```{r}
summary(dataWork)
```

## Check Observed and Simulated time-series

- Show time-series for dryland crops

- Simulated parameter combinations' data (grey) and observed values (red)

- Note that, in some cases, observations are at the edge of simulation's range. 

- <span style="color:red">To discuss: Is this a sign that parameter ranges should be expanded???</span>  

```{r GraphExample, fig.height=18,fig.width=12, echo=FALSE}

# example of dryland crops
dataWork %>%
  filter(Irrigation == "No Irrigation") %>%
  ggplot() +
  geom_point(aes(x=Date, y=Pred),colour="grey") +
  geom_point(aes(x=Date, y=Obs), size=2, colour="darkred") +
  facet_grid(Depth~Crop) +
  geom_vline(xintercept = as.numeric(as.Date(ymd(estDate))), linetype=2) 

```

## Graph Predicted x Observed

- For whole dataset (all parameter combinations)
- 1:1  line in black (x=y)
- linear regression for each soil depth shown in gray
- 95% confidence interval also shown

```{r XYgraph , fig.height=18,fig.width=12, echo=FALSE}
dataWork %>%
  ggplot(aes(x=Pred, y=Obs,colour=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
              linetype = 3,
              aes(colour=factor(Depth))) +
  geom_point() +
  facet_grid(Crop~Stage, scales ="free") +
  geom_abline(intercept = 0, slope = 1) + 
  coord_fixed(ratio = 1) +
  xlim(0,max(dataWork$Pred,dataWork$Obs)) +
  ylim(0,max(dataWork$Pred,dataWork$Obs))

```

## Create Gauch stats function

- Function to calculate RMSE components
- Standard bias (SB)
- Non unit (NU)
- Lack of correlation (LC)
- Reference: Gauch Jr, H.G., Hwang, J.T.G., Fick, G.W., 2003. Model evaluation by comparison of model-based predictions and measured values. Agron. J. 95, 1442-1446

```{r CreateStats, echo=FALSE, include=FALSE}

# Create stats function as per Gauch et al paper

gauchStats <- function(sim, meas) {

  n_s <- length(sim)
  n_m <- length(meas)
  model <- lm(meas~sim)
  sim_sq <- sum((sim - mean(sim))^2)
  mes_sq <- sum((meas - mean(meas))^2)
  r2 <- summary(model)$r.squared
  slope <- model$coefficients[[2]]

  sb <- (sum(mean(meas)) - sum(mean(sim)))^2
  nu <- (1-slope)^2 * (sim_sq/n_s)
  lc <- (1-r2) * (mes_sq/n_m)
  msd <- sb+nu+lc

  sb_r <- round((sb/msd)*100,1)
  nu_r <- round((nu/msd)*100,1)
  lc_r <- round((lc/msd)*100,1)

  msd_r <- sb_r+nu_r+lc_r

  # select which variables to output
  out <- c(sb_r, # output 1
           nu_r,
           lc_r, 
           msd_r, 
           round(r2*100,1)) # output 5

  return(out)

}



```

- Test Gauch stats
- Result has to be: 76.5  5.2 18.2 99.9  7.1

```{r, include=FALSE}
# test dataset
s <- c(4231.972,3935.604,3779.652,3627.687,3363.499,3230.566,2868.114,2868.827)
m <- c(4987.66,5636.09,4754.06,4114.53,4141.72,3704.06,5142.19,4762.03)

gauchStats(s,m)
```

## Creates a generic set of stats to output 

- These are the standard stats used in paper
- Basic stats (e.g. r2, RMSE ...) plus the Gauch et al. components defined in the function above

```{r, echo = FALSE, include = TRUE}

# Create a tailored common function that do the stats and includes the gauch functionality
# input is a df with Pred and Obs columns

doStats <-  function (x) {
    library(dplyr)
    x %>% dplyr::summarise(
    n = n(),
    r2 = gauchStats(Pred,Obs)[5],
    rmse = rmse(Pred,Obs),
    r_rmse = rmse(Pred,Obs)/mean(Obs)*100, # big precision to enable cretrial of selection
    nse = NSE(Pred,Obs),
    sb = gauchStats(Pred,Obs)[1],
    nu = gauchStats(Pred,Obs)[2],
    lc = gauchStats(Pred,Obs)[3]) 
}

```

## Create Indexes for data aggregation for stats

- Aggregate Pred x Obs in different forms to apply stats
- The issue is to define what's the creteria to define the best fit
- For example, to test parameter combinations for "each" Crop and "each" Stage we agregate by "Crop and Stage"
- This e.g. then calculates stats across "all" soil depths and "all" water-treatments.
- Multiple agregations are done below

```{r CreateIndexes}

# Best fit by crop only ---------------------------------------
dataWork$index_crop <- paste0(dataWork$Crop, "_SK",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop <- as.factor(dataWork$index_crop)


# Best fit  by crop and "stage" ---------------------------------------------
dataWork$index_crop_stage <- paste0(dataWork$Crop, "_",
                         dataWork$Stage, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_stage <- as.factor(dataWork$index_crop_stage)

# Best fit  by crop and "depth" ---------------------------------------
dataWork$index_crop_depth <- paste0(dataWork$Crop, "_",
                              dataWork$Depth, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_depth <- as.factor(dataWork$index_crop_depth)


# Best fit  by crop, stage and depth ---------------------------------------
dataWork$index_crop_stage_depth <- paste0(dataWork$Crop, "_",
                              dataWork$Stage, "_Dp:",
                              dataWork$Depth, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_stage_depth <- as.factor(dataWork$index_crop_stage_depth)

# re-order fators for tidy up (ensure Obs is the LAST one)
dataWork <- dataWork[c("Crop", "Irrigation",            
                       "Depth",
                       "Stage",
                       "Date",
                       "SurfaceKl", 
                       "KLReduction", 
                       "RootFrontVelocity",
                       "index_crop",
                       "index_crop_stage",
                       "index_crop_depth",
                       "index_crop_stage_depth",
                       "Pred", "Obs")]
```

## Calculate stats

- Do calculations for each agregation possibility

- By crop

```{r DoStatsCrop, echo=TRUE, include=TRUE}

# Separate Crop only
statDF_cropOnly <- dataWork  %>%
group_by(Crop, SurfaceKl, KLReduction, RootFrontVelocity, index_crop) %>%
doStats() 

summary(statDF_cropOnly)
```

- By crop & Stage

```{r}
# Separate Crop & Stage
statDF_cropStage <- dataWork %>%
  dplyr::group_by(Crop, Stage, SurfaceKl, KLReduction, RootFrontVelocity, index_crop_stage) %>%
  doStats()

summary(statDF_cropStage)
  
```

- Crop & Depth

```{r}

# Separate by Crop & Depth
statDF_cropDepth <- dataWork %>%
  dplyr::group_by(Crop, Depth, SurfaceKl, KLReduction, RootFrontVelocity, 
                  index_crop,
                  index_crop_depth) %>%
  doStats() 

summary(statDF_cropDepth)

```

- By crop & stage & depth

```{r}
# Separate Crop & Stage & Depth
statDF_cropStageDepth <- dataWork %>%
  dplyr::group_by(Crop, Stage, Depth, SurfaceKl, KLReduction, RootFrontVelocity, 
                  index_crop,   # has to keep other indexes as we'll use them later
                  index_crop_stage, 
                  index_crop_stage_depth) %>%
  doStats() 

summary(statDF_cropStageDepth)

```

## Find and exclude "bad" parameter combinations at individual soil depths

- Aim: avoid overall good fit but bad fit(s) in individual soil depth layers

- Find and excludes parameter combinations that give poor fit in individual layers (i.e. by Depth)

- Start with Crop x Stage x Depth combinations (i.e. statDF_cropStageDepth index)

- Criteria of exclusion: r_rmse of layer is > upper quartile r_rmse of simulation population

- <span style="color:red">Is this necessary and a good criteria?</span>

```{r ExcludeBadLayer}

# FIXME: there's a duplication of this selection below by calculating 25% percentiles (low rmse) need to decide and keep only one of them. This one is at "soil depth layer" level though.

# Find the 75% percentile for each crop. 
# obs: Upper quartile is the 4th vector element of the quantile function output
cutByLayer_luc <- quantile(statDF_cropStageDepth$r_rmse[statDF_cropStageDepth$Crop == "Lucerne"])[[4]] # lucerne
cutByLayer_pas <- quantile(statDF_cropStageDepth$r_rmse[statDF_cropStageDepth$Crop == "Pasture"])[[4]] # pasture

print(paste0("Lucerne upper quartile r_rmse: ",round(cutByLayer_luc,1),"% mean"))

print(paste0("Pasture upper quartile r_rmse: ",round(cutByLayer_pas,1),"% mean"))

```

- Simulations to be excluded due to having one or more poor soil layers being simulated

```{r, include=FALSE, echo = FALSE}

# agregated by crop only - NOT USED

# Find & exclude any parameter combination simulations with "individual" soil depth layers with r_rmse > upper quartile of r_rmse
remove_badLayer_byCrop <- statDF_cropStageDepth %>%
  rowwise() %>%
  filter(ifelse(Crop == "Lucerne", 
                r_rmse>cutByLayer_luc, 
                r_rmse>cutByLayer_pas)) %>%
  dplyr::select(index_crop, Crop, Depth, r_rmse) %>%
  unique()

summary(remove_badLayer_byCrop)
```


```{r, include=FALSE, echo = FALSE}

# agregated by crop only - NOT USED

# - Graphs show the number of "excluded" simulations at a given depth

# - Treat crops separatelly but excludes bad parameter combinations from both stages (Establishment and Regrowth)

remove_badLayer_byCrop %>%
  mutate(Depth = as.numeric(as.character(Depth)))  %>%
  ggplot(aes(x=Depth)) +
  geom_histogram(aes(y=..count..)) +
  facet_grid(.~Crop)

```

- Identify soil layer with bad fit for each "crop by stage" combinations"

- <span style="color:red">Is this rationale justifiable or am I over-doing?</span>

```{r}

# By crop and stage
remove_badLayer_byCropStage <- statDF_cropStageDepth %>%
  rowwise() %>%
  filter(ifelse(Crop == "Lucerne", r_rmse>cutByLayer_luc, r_rmse>cutByLayer_pas)) %>%
  dplyr::select(index_crop_stage, Crop, Stage, Depth) %>%
  unique()

#summary(statDF_cropStageDepth)
#summary(remove_badLayer_byCropStage)

all_df <- statDF_cropStageDepth %>% group_by(Crop, Stage) %>% summarise(all = n())
remove_df <- remove_badLayer_byCropStage %>% group_by(Crop, Stage) %>% summarise(remove = n())

df <- merge(all_df, remove_df, by = c("Crop","Stage"))
df <- df %>% mutate(remove_Perc = round(remove/all*100,1)) # percentage of excluded simulations with at least one layer with bad fit

head(df)


```

- Graph number of simulations removed by poor fit in any given soil layer
- Note that poor fit in individual layers was most frequently observed in regrowth stage
- <span style="color:red">These parameter combinations will be excluded</span>
- <span style="color:red">FIXME: How do I know if good overall sims are not excluded unfaily here?</span>

```{r}

# FIXME: this will happen as last step after rmse/LC optimisation - to be moved
remove_badLayer_byCropStage %>%
  mutate(Depth = as.numeric(as.character(Depth)))  %>%
  ggplot(aes(x=Depth)) +
  geom_histogram(aes(y=..count..)) +
  facet_grid(Stage~Crop)
```

## Compare stats used

- Aim: Decide which stats to use as the criteria to select "optimum" parameter combinations
- Note that many stats are "highly" correlated (e.g. see nse x r_rmse x rmse)


```{r AnalyseStats, fig.height=10,fig.width=12, echo=FALSE, warning=FALSE}

# relationship between stats (Obs: it crashes if try to plot Crop+Stage or Crop+Stage+Depth - too heavy)
# plot(statDF_cropOnly[,7:13], col=statDF_cropOnly$Crop)
# legend(7,4.3,unique(statDF_cropOnly$Crop),col=1:length(statDF_cropOnly$Crop),pch=1)

# with stats by Crop
statDF_cropOnly %>%
 group_by(Crop) %>%
dplyr::select(r2:lc) %>%
ggpairs(aes(colour=Crop),labelHJust=1) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0))

# ggpairs(statDF_cropOnly[,7:13], colour = statDF_cropOnly$Crop)

```

- So the idea is to combine stats that have low correlation with each other

- At the moment I have chosen:

- <span style="color:red">minimise r_rmse</span> : low error 

- <span style="color:red">maximise LC</span> : low bias

- <span style="color:red">Is this a good rationale? Alternatives?</span>


## Select similations with good fit

- Use stats done by Crop and Stage
- Option 1: filter out simulations with bad fit in individual layers <span style="color:red">NOT used at the moment</span>
- Option 2: keep all simulations <span style="color:red">Used at the moment</span>
- Check which one is being used

```{r}

vcTemp <- remove_badLayer_byCropStage$index_crop_stage

# Excludes sims with soil layers with bad fit
 df_selected <- statDF_cropStage %>% 
   filter(!(index_crop_stage %in% vcTemp )) # FIXME: do this later - if done now it excludes too many 

nrow(df_selected)

# Keep all parameter combinations
df_selected <- statDF_cropStage

nrow(df_selected)

```

## Map stats for all parameter combinations

- Plot r_rmse values
- Blue is low r_rmse (good) and red is high r_rmse (bad)
- Apologies for the horrible colours (FIXME)

```{r, fig.height=8,fig.width=8, echo=FALSE}

# Create categories of r_rmse (to see colour differences)
df_selected$stats_cat <- cut(df_selected$r_rmse, breaks = 10) # select stats here
lv <- length(unique(df_selected$stats_cat))
colors <- colorRampPalette(c("blue", "green", "yellow", "red"))(lv)
df_selected$stats_cat <- factor(df_selected$stats_cat)
N <- nlevels(df_selected$stats_cat)

colfunc <- colorRampPalette(c("blue", "green", "yellow", "red"))

# Map lc
df_selected  %>% 
  ungroup() %>% # to allow mutate
  group_by(Crop, Stage) %>%
  ggplot(aes(x = SurfaceKl, 
             y = KLReduction, 
             z = RootFrontVelocity, 
             fill = stats_cat)) +
  geom_tile() + 
  coord_equal() +
  stat_contour() + 
  facet_grid(Stage~Crop+RootFrontVelocity) + 
 # scale_fill_manual(values=colors, breaks=levels(statDF_cropStage$stats_cat)[seq(1, N, by=1)])+
  scale_fill_manual(values=colfunc(length(levels(df_selected$stats_cat)))) +
  xlab("Surface kl category") +
  ylab("Exponential kl reduction category")+
  theme(text = element_text(size=16)) # + theme(axis.text.x = element_text(angle = 90, hjust = -0.5))

```

- Plot LC values
- Blue is high LC (good) and red is low LC (bad)


### Ideas to ncorporate:

- do layer selection (exclusion) later after the overall fit was selected
- plot a semnsitivity graph to different indexes (done already?)

```{r, fig.height=8,fig.width=8, echo=FALSE, warning=FALSE}

# Create categories of r_rmse (to see colour differences)
df_selected$stats_cat <- cut(df_selected$lc, breaks = 10) # select stats here
lv <- length(unique(df_selected$stats_cat))
colors <- colorRampPalette(c("blue", "green", "yellow", "red"))(lv)
df_selected$stats_cat <- factor(df_selected$stats_cat)
N <- nlevels(df_selected$stats_cat)

colfunc <- colorRampPalette(c("blue", "green", "yellow", "red"))

# Map lc
df_selected  %>% 
  ungroup() %>% # to allow mutate
  group_by(Crop, Stage) %>%
  ggplot(aes(x = SurfaceKl, 
             y = KLReduction, 
             z = RootFrontVelocity, 
             fill = stats_cat)) +
  geom_tile() + 
  coord_equal() +
  stat_contour() + 
  facet_grid(Stage~Crop+RootFrontVelocity) + 
 # scale_fill_manual(values=rev(colors), breaks=levels(statDF_cropStage$stats_cat)[seq(1, N, by=1)])+
  scale_fill_manual(values=colfunc(length(levels(df_selected$stats_cat))))+
  xlab("Surface kl category") +
  ylab("Exponential kl reduction category")+
  theme(text = element_text(size=16)) # + theme(axis.text.x = element_text(angle = 90, hjust = -0.5))
```

## Distribution of r_rmse vales for each parameter

```{r, fig.height=8,fig.width=8, echo=FALSE, include = FALSE}

# Found harder to evaluate distributions for this case (Not used)

# KLReduction
df_selected %>%
  ungroup() %>% # to allow mutate
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=r_rmse, colour=factor(RootFrontVelocity),fill=factor(RootFrontVelocity))) +
  geom_density(alpha=0.2) + 
  facet_grid(Stage~Crop)

# SurfaceKl
df_selected %>%
  ungroup() %>% # to allow mutate
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=r_rmse, colour=factor(SurfaceKl),fill=factor(SurfaceKl))) +
  geom_density(alpha=0.2) + 
  facet_grid(Stage~Crop)

# KLReduction
df_selected %>%
  ungroup() %>% # to allow mutate
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=r_rmse, colour=factor(KLReduction),fill=factor(KLReduction))) +
  geom_density(alpha=0.2) + 
  facet_grid(Stage~Crop)



```

### RootFrontVelocity (RFV) parameter

- RFV has more leverage during Establishment (as expected)
- RFV of 5 gives poor fit but has lowest bias (high LC) - trade-off!
- RFV of 15 seems good for both crops
- <span style="color:red">How to select given the trade-off between LC and r_RMSE?</span>

```{r, fig.height=7,fig.width=10, echo=FALSE, include = TRUE}

df_selected %>%
  ungroup() %>% # to allow mutate
#  mutate(lc_cat = cut(lc, breaks = quantile(lc, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  tidyr::gather("StatType", "StatValue", n:lc) %>%
  filter(StatType == "r_rmse" | StatType == "lc") %>%
  mutate(StatType = factor(StatType, levels = c("r_rmse","lc"))) %>%
  ggplot(aes(x=factor(RootFrontVelocity), y=StatValue , colour=factor(RootFrontVelocity))) +
  geom_jitter(alpha=0.2,width=0.2) +
  geom_boxplot(alpha=0.2) + 
  facet_wrap(Crop~StatType+Stage, scales="free",ncol = 4)

```

### Surface kl parameter


```{r, fig.height=7,fig.width=10, echo=FALSE, include = TRUE}
df_selected %>%
  ungroup() %>% # to allow mutate
  tidyr::gather("StatType", "StatValue", n:lc) %>%
  filter(StatType == "r_rmse" | StatType == "lc") %>%
  mutate(StatType = factor(StatType, levels = c("r_rmse","lc"))) %>%
  ggplot(aes(x=factor(SurfaceKl), y=StatValue , colour=factor(SurfaceKl))) +
  geom_jitter(alpha=0.2,width=0.2) +
  geom_boxplot(alpha=0.2) + 
  facet_wrap(Crop~StatType+Stage, scales="free",ncol = 4)
```


```{r, include =FALSE}

# chunck to check

df_selected %>%
  ungroup() %>% # to allow mutate
  mutate(lc_cat = cut(lc, breaks = quantile(lc, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(KLReduction), y=lc , colour=factor(KLReduction))) +
  geom_jitter(aes(size= factor(KLReduction)), width=0.2,alpha=0.2)+
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop, scales="free") +
 theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Sensitivity plots for lc and rmse
df_selected %>%
  ungroup() %>% # to allow mutate
  mutate(lc_cat = cut(lc, breaks = quantile(lc, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(SurfaceKl), y=r_rmse , colour=factor(SurfaceKl))) +
  geom_jitter(aes(size= factor(KLReduction)), width=0.2,alpha=0.2)+
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop, scales="free") 


# Sensitivity plots for lc and rmse
df_selected %>%
  ungroup() %>% # to allow mutate
  mutate(lc_cat = cut(lc, breaks = quantile(lc, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(SurfaceKl), y=lc , colour=factor(SurfaceKl))) +
  geom_jitter(aes(size= factor(KLReduction)), width=0.2,alpha=0.2)+
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop, scales="free") 


```



```{r, include=FALSE}

# chunck to check

# Sensitivity plots for lc and rmse
df_selected %>%
  ungroup() %>% # to allow mutate
  mutate(r_rmse_cat = cut(r_rmse, breaks = quantile(r_rmse, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(SurfaceKl), y=lc , colour=factor(SurfaceKl))) +
  geom_jitter(aes(size= factor(r_rmse_cat)), width=0.2,alpha=0.2)+
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop, scales="free") 

# RFV

# Sensitivity plots for lc and rmse
df_selected %>%
  ungroup() %>% # to allow mutate
  mutate(lc_cat = cut(lc, breaks = quantile(lc, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(RootFrontVelocity), y=r_rmse , colour=factor(RootFrontVelocity))) +
  geom_jitter(aes(size= factor(lc_cat)), width=0.2,alpha=0.1)+
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop, scales="free") 

# Sensitivity plots for lc and rmse
df_selected %>%
  ungroup() %>% # to allow mutate
  mutate(r_rmse_cat = cut(r_rmse, breaks = quantile(r_rmse, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(RootFrontVelocity), y=lc , colour=factor(RootFrontVelocity))) +
  geom_jitter(aes(size= factor(r_rmse_cat)), width=0.2,alpha=0.1)+
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop, scales="free") 

# KL reduction

# Sensitivity plots for lc and rmse
df_selected %>%
  ungroup() %>% # to allow mutate
  mutate(lc_cat = cut(lc, breaks = quantile(lc, probs = seq(0, 1, 0.2)))) %>%
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(RootFrontVelocity), y=r_rmse , colour=factor(RootFrontVelocity))) +
  geom_jitter(aes(size= factor(lc_cat)), width=0.2,alpha=0.1)+
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop, scales="free") 


```



```{r}

# Sensitivity plots for lc and rmse
df_selected %>%
  ungroup() %>% # to allow mutate
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=factor(KLReduction), y=r_rmse , colour=factor(KLReduction))) +
  geom_boxplot(alpha=0.2) + 
  facet_grid(Stage~Crop) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}

head(df_selected)

```

# Calculates the best fit (lowest r_rmse)

## For the "establishment phase only"

- Searches for best fit combination in establishment phase
- Selects root front velocity for that phase

```{r BestFitFinder1, echo=FALSE}
# FIXME: Finds selected qualtile threshold of RMSE (i.e. lowest X% of RMSE values) - not used yet - or is it?

 # Finds the 25% best fit (i.e. lowest r_rmse)
 p <- 0.25
 Q_luc_est <- df_selected %>% subset(Crop == "Lucerne" & Stage == "Establishment") %>% mutate(Q = quantile(rmse,probs=p))
 Q_luc_reg <- df_selected %>% subset(Crop == "Lucerne" & Stage == "Regrowth") %>% mutate(Q = quantile(rmse,probs=p))  
 Q_pas_est <- df_selected %>% subset(Crop == "Pasture" & Stage == "Establishment") %>% mutate(Q = quantile(rmse,probs=p)) 
 Q_pas_reg <- df_selected %>% subset(Crop == "Pasture" & Stage == "Regrowth") %>% mutate(Q = quantile(rmse,probs=p))  

 # FIXME: does not calculate quartile per group
 # df_selected %>% group_by(Crop,Stage) %>% mutate(Q = quantile(rmse,probs=p)) %>% summary() 

#Find good fit 25% population quartiles for each crop/stage 
# FIXME: Sure there's a way to do this in groups (not used)
# FIXME: this overwrites df_selected!
df_selected <- df_selected %>% mutate(rmseCut = ifelse(grepl("Lucerne_Est", index_crop_stage), Q_luc_est$Q[1],
                                ifelse(grepl("Lucerne_Reg", index_crop_stage), Q_luc_reg$Q[1],
                                    ifelse(grepl("Pasture_Est", index_crop_stage), Q_pas_est$Q[1],
                                       ifelse (grepl("Pasture_Reg", index_crop_stage), Q_pas_reg$Q[1], NA)))
                                ))
summary(df_selected)

```

```{r, fig.height=12,fig.width=12, echo=FALSE, warning=FALSE}

# finds lowest r_rmse for the establishment phase first
bestFit_Estab <- df_selected  %>%
  filter(Stage == "Establishment") %>%
  group_by(Crop) %>%
  filter(rmse<=rmseCut) %>% # was selecting good fit population
#  filter(r_rmse == min(r_rmse)) %>%
  filter(lc == max(lc)) %>%
  filter(r_rmse == min(r_rmse) )

rownames(bestFit_Estab) <- paste0(bestFit_Estab$Crop)

t(bestFit_Estab) %>%
  kable(format = "markdown")

selecParamsEst <- as.character(unique(bestFit_Estab$index_crop_stage))
RVF_luc_est <- bestFit_Estab[bestFit_Estab$Crop=="Lucerne",]$RootFrontVelocity
RVF_rye_est <- bestFit_Estab[bestFit_Estab$Crop=="Pasture",]$RootFrontVelocity

```

## Select best fit for Crop and Stages

- Uses "only" Root Front Velocity found in "establishment" phase (i.e. filters out other RFVs)

```{r, fig.height=12,fig.width=12, echo=FALSE, warning=FALSE}

# now select best fit for regrowth by setting the RFV to the selected in the establishment phase
bestFit_Reg <- df_selected %>% 
  filter(  
    index_crop_stage == selecParamsEst[1] | # bring back only selected establishmnet combinations 
    index_crop_stage == selecParamsEst[2] |
    (Crop == "Lucerne" &  Stage == "Regrowth" & RootFrontVelocity == RVF_luc_est) |   
    (Crop == "Pasture" &  Stage == "Regrowth" & RootFrontVelocity == RVF_rye_est)) %>% 
  group_by(Crop, Stage) %>%
  filter(rmse<=rmseCut) %>%
 # filter(r_rmse == min(r_rmse)) %>%
  filter(lc == max(lc)) %>%
  filter(r_rmse == min(r_rmse) )
  
rownames(bestFit_Reg) <- paste0(bestFit_Reg$Crop,".",bestFit_Reg$Stage)

t(bestFit_Reg) %>%
  kable(format = "markdown")

```

## Best fit graphs for LUCERNE

- Red circles are observed
- Black line is "best" fit
- Grey circles are all simulations (i.e. parameter combinations)

```{r PlotBestFitsLuc, fig.height=18,fig.width=12, echo=FALSE}

selecParams <- as.character(unique(bestFit_Reg$index_crop_stage))

sel1 <- dataWork %>%
  filter(
    # establisment
    index_crop_stage == selecParams[1] | # watch for this conventions of names (FIXME)
    # regrowth phase
    index_crop_stage == selecParams[2]
  ) 

# finds the closest point to the date of establihment
estLoc <- which.min(abs(as.Date(sel1$Date) - as.Date(estDate)))

# graph lucerne
sel1 %>%
  ggplot() +
  geom_point(data=dataWork[dataWork$Crop == "Lucerne",], aes(x=Date, y=Pred), colour="grey",alpha=0.1) +
  geom_point(aes(x=Date, y=Obs), size=2, shape = 1, colour="darkred") +
  geom_line(aes(x=Date, y=Pred),colour="black") +
  facet_grid(Depth~Irrigation, scales = "free") + 
   facet_grid(Depth~Irrigation) + 
  scale_x_date(labels = date_format("%b")) + scale_shape(solid = FALSE) + 
  geom_vline(xintercept = as.numeric(sel1$Date[estLoc]),linetype = 2) +
  ylab("Volumetric soil moisture (mm)") +
  theme(text = element_text(size=16)) 

#sel1  %>%
#  ggplot(aes(x=Pred, y=Obs,fill = factor(Stage),colour=factor(Stage))) +
#  stat_smooth(method = "lm", se = TRUE, 
#            linetype = 3) +
#  geom_point() +
#  facet_grid(Depth~Irrigation, scales = "free") +
#  geom_abline(intercept = 0, slope = 1) 

# Idea: plot range of simulations around best fit 

```

```{r, fig.height=8,fig.width=8, echo=FALSE}
sel1  %>%
  ggplot(aes(x=Pred, y=Obs,fill = factor(Depth),colour=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
            linetype = 3) +
  geom_point() +
  facet_grid(Stage~Irrigation, scales = "free") +
  geom_abline(intercept = 0, slope = 1) +
  ylab("Observed") +
  xlab("Predicted")+ 
  coord_fixed(ratio = 1) +
  xlim(0,max(sel1$Pred,sel1$Obs)) +
  ylim(0,max(sel1$Pred,sel1$Obs))
```

## Best fit graphs for RYEGRASS

```{r BestFitRye, fig.height=8,fig.width=8, echo=FALSE}

# Ryegrass
sel2 <- dataWork %>%
  filter(
    # establisment
    index_crop_stage == selecParams[3] |
    # regrowth phase
    index_crop_stage == selecParams[4]
  ) 


# graph ryegrass
 sel2 %>%
  ggplot() +
  geom_point(data=dataWork[dataWork$Crop == "Pasture",], aes(x=Date, y=Pred), colour="grey",alpha=0.1) + 
  geom_point(aes(x=Date, y=Obs), size=2, shape = 1, colour="darkred") +
  geom_line(aes(x=Date, y=Pred),colour="black") +
  facet_grid(Depth~Irrigation) + 
  scale_x_date(labels = date_format("%b")) + scale_shape(solid = FALSE) + 
  geom_vline(xintercept = as.numeric(sel1$Date[estLoc]),linetype = 2)+
  ylab("Volumetric soil moisture (mm)") +
  theme(text = element_text(size=16)) 

#sel2  %>%
#  ggplot(aes(x=Pred, y=Obs,fill = factor(Stage),colour=factor(Stage))) +
#  stat_smooth(method = "lm", se = TRUE, 
#            linetype = 3) +
#  geom_point() +
#  facet_grid(Depth~Irrigation, scales = "free") +
#  geom_abline(intercept = 0, slope = 1) 

               



```

```{r, fig.height=12,fig.width=12, echo=FALSE}
sel2  %>%
  ggplot(aes(x=Pred, y=Obs,fill = factor(Depth),colour=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
            linetype = 3) +
  geom_point() +
  facet_grid(Stage~Irrigation, scales = "free") +
  geom_abline(intercept = 0, slope = 1)  +
  ylab("Observed") +
  xlab("Predicted") +
  xlim(0,max(sel2$Pred,sel1$Obs)) +
  ylim(0,max(sel2$Pred,sel1$Obs))
```

## How does the best fit(s) look for RMSE by depth?

```{r RmsdPerDepth, echo = FALSE, fig.height=12,fig.width=12, echo=FALSE}

# calculate the stats for the BEST fit
# For each layer and depth combination to show trends
# IDEA: how much this afected water uptake?

sel3 <- rbind(sel1, sel2)

# calculates stats by depth
g1 <- sel3 %>%
  dplyr::group_by(Crop, Stage, Irrigation, Depth, Depth, SurfaceKl, KLReduction, RootFrontVelocity) %>%
  dplyr::summarise(
    n = n(),
    r2 = gauchStats(Pred,Obs)[5],
    rmse = round(rmse(Pred,Obs),3),
    r_rmse = round(rmse(Pred,Obs)/mean(Obs)*100,3),
    nse = round(NSE(Pred,Obs),1),
    sb = gauchStats(Pred,Obs)[1],
  nu = gauchStats(Pred,Obs)[2],
  lc = gauchStats(Pred,Obs)[3]) 

head(g1,20)



```

```{r, fig.height=12,fig.width=12, echo=FALSE, include=FALSE}

# this graph is not adding to previous ones, therefore not used

# create a factor of  depth as integer
#g1$Depth <- as.numeric(gsub("D", "",g1$Depth)) # FIXME: move this to a higher level (earlier) as it can be used in other graphs

# _rmse by depth and crop
g1 %>%
  ggplot(aes(x=as.numeric(as.character(Depth))*-1, y=r_rmse, colour=factor(Stage))) +
  geom_point() +
  geom_line() +
 # geom_bar(aes(stats='identity')) +
  facet_grid(Crop~Irrigation) +
  coord_flip() + 
  geom_hline(yintercept = 20,linetype = 20) +
  labs(x = "Soil depth (cm)", y = "Relative Root Mean Squared Error (RMSE, % mean)") +
  theme(text = element_text(size=rel(4))) +
  theme(legend.text=element_text(size=16))
```


## RMSE components per depth

- Compare stats across depths now

- Partition of RMSE into Gauch fractions

- Standard bias (SB)

- Non Unit (NU)

- Lack of Correlation (LU)

```{r GauchStatsByDepth, echo = TRUE, fig.height=8,fig.width=8, echo=FALSE, warning=FALSE}

df <- g1 %>%
  mutate(mse = rmse^2, sb=sb*mse, nu=nu*mse, lc=lc*mse) %>%
 # dplyr::select(-n, -r2,-rmse,-r_rmse,-nse) %>%
  dplyr::select(-n, -r2,-r_rmse,-nse, -rmse) %>%
  tidyr::gather("StatsType","StatsValue", 8:10) %>%
  mutate(StatsType = factor(StatsType, levels = c("sb","nu","lc", "mse")))

# graph
  df %>%
#  filter(Crop == "Lucerne") %>%
  ungroup() %>%  
  mutate(Depth = factor(Depth, level = rev(Depth))) %>%
  ggplot(aes(x=Depth, y=StatsValue,fill=rev(StatsType))) +
  geom_bar(stat='identity') +
  facet_grid(Irrigation~Crop+Stage,scales = "free") +
  coord_flip() + 
  labs(x = "Soil depth (cm)", y = "MSE") +
  theme(text = element_text(size=rel(4))) +
  theme(legend.text=element_text(size=16))

```

```{r PinPointBestFit, echo = FALSE, fig.height=12,fig.width=12, include=FALSE}

# create a circle around the best fits
circleFun <- function(center = c(0,0),diameter = 1, npoints = 100){
    r = diameter / 2
    tt <- seq(0,2*pi,length.out = npoints)
    xx <- center[1] + r * cos(tt)
    yy <- center[2] + r * sin(tt)
    return(data.frame(x = xx, y = yy))
}

dat <- circleFun(c(50,20),4, npoints = 100)
#geom_path will do open circles, geom_polygon will do filled circles
ggplot(dat,aes(x,y)) + geom_path()

# 
# dfTemp <- statDF_cropStage %>%
#   group_by(Crop, Stage) %>%
#   dplyr::select(lc, r_rmse) %>%
#   summarise_each(funs(max, min)) %>%
#   dplyr::select(Crop, Stage, lc_max,r_rmse_min)







```

## Plot population of simulations 

- uses rRMSE and LC as criteria

```{r}



# Flag simulations that gave best fit
temp_data_bf <- statDF_cropStage %>%
  mutate(BestFit = ifelse(index_crop_stage %in% bestFit_Reg$index_crop_stage,"Best", "NotBest")) %>%
  mutate(BestFit = factor(BestFit, levels=c("NotBest", "Best"))) 

#summary(temp_data_bf)

# create df with "best fit" stat values and ranges
rangeBestFit <- 0.05 # % of value

bound_bf <- temp_data_bf %>%
  filter(BestFit == "Best") %>%
  group_by(Crop, Stage) %>%
  dplyr::select(lc, r_rmse) %>%
  mutate(r_rmse_best = r_rmse, lc_best = lc) %>%
  dplyr::select(-lc, -r_rmse) %>%
  mutate(r_rmse_bestFit_low = (1-rangeBestFit)*r_rmse_best,
         r_rmse_bestFit_up = (1+rangeBestFit)*r_rmse_best,
         lc_bestFit_low = (1-rangeBestFit)*lc_best,
         lc_bestFit_up = (1+rangeBestFit)*lc_best)

summary(bound_bf)

# Create a flag in big dataset to point if r_rmse and lc are within range of good fit
tempdf <- merge(temp_data_bf,bound_bf, by=c("Crop", "Stage"))

data_bf <- tempdf %>%
  rowwise() %>%
  mutate(r_rmse_range_flag = ifelse(r_rmse < r_rmse_bestFit_up, 1, 0)) %>%
  mutate(lc_range_flag = ifelse(lc <lc_bestFit_low, 0, 1)) %>%
  mutate(range_idx = r_rmse_range_flag * lc_range_flag) %>%
  dplyr::select(-r_rmse_range_flag,-lc_range_flag,-(r_rmse_bestFit_low:lc_bestFit_up))


# range_bf_df <- data_bf %>%
#   group_by(Crop, Stage) %>%
#   filter(BestFit == "Best") %>%
#   mutate(r_rmse_bestFit_low = (1-rangeBestFit)*r_rmse,
#          r_rmse_bestFit_up = (1+rangeBestFit)*r_rmse,
#          lc_bestFit_low = (1-rangeBestFit)*lc,
#          lc_bestFit_up = (1+rangeBestFit)*lc)


summary(data_bf)



```

```{r, fig.height=12,fig.width=12, echo=FALSE}
# cbPalette <- c("#999999", "#E69F00","#56B4E9", "#009E73","#999999", "#E69F00","#56B4E9", "#009E73")
#data_bf %>%
#  filter(range_idx > 0)

data_bf %>%
  ggplot(aes(y=lc, x=r_rmse)) + # not working yet
  geom_point(aes(size=factor(SurfaceKl),
                 colour=factor(KLReduction),
                 shape=factor(RootFrontVelocity))) +
#  geom_path(aes(x,y), data=dat) +
  stat_ellipse(type = "norm",aes(linetype = range_idx == 1)) + 
  scale_linetype_manual(values=c("blank", "dotted")) +
  scale_shape_manual(values=c(21,22,24))  +
  facet_wrap(Crop~Stage, scales= "free") +
  ylab("Lack of correlation (LC, % MSE) ") +
  xlab(" Relative RMSE (% observed mean)") +
  geom_vline(data=bound_bf, aes(xintercept=r_rmse_best), linetype=2) + #
  geom_hline(data=bound_bf, aes(yintercept=lc_best), linetype=2) # 
  
  # geom_hline(data=dfTemp, aes(yintercept=r_rmse_min), linetype=2) +
  # geom_vline(data=dfTemp, aes(xintercept=lc_max), linetype=2)

# idea: find values around 10% of RMSE and 10% of LC and draw ellipse

```


```{r}



data_bf %>%
 filter(range_idx == 1) %>%
  ggplot(aes(x=as.numeric(as.character(SurfaceKl)),y= as.numeric(as.character(RootFrontVelocity)))) +
    geom_count() +
    facet_wrap(Stage~Crop, scales= "free") 

data_bf %>%
 filter(range_idx == 1) %>%
  ggplot(aes(x=as.numeric(as.character(KLReduction)),y= as.numeric(as.character(RootFrontVelocity)))) +
    geom_count() +
    facet_wrap(Stage~Crop, scales= "free")

```

```{r, fig.height=12,fig.width=12, echo=FALSE}
data_bf %>%
 filter(range_idx == 1) %>%
  group_by(SurfaceKl,KLReduction) %>%
  ggplot(aes(x=as.numeric(as.character(SurfaceKl)),y= as.numeric(as.character(KLReduction)))) +
  # geom_count() +
    geom_point(aes(size=rmse)) +
    facet_wrap(Stage~Crop+RootFrontVelocity, scales= "free") 
```

## 

- From the best fit ...
- Check individual layers and keep only best 

```{r}




```


## Ideas to explore

- Do the ANOVA sensitivity analysis to find out which parameters are the most important
