---
title: "Analysis of L&P dataset"
output: html_document
---


```{r LoadLibs, echo=FALSE, include=FALSE}
library(xtable)
library(knitr)
library(RSQLite)
library(lubridate)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library (hydroGOF)
library(scales)

options("scipen"=100, "digits"=4) # prints not in scientific notation

```


## FIXME list

- Print numeric value of Surfacekl in csv output
- Use numeric value of parameters as characters througout to simplify comparisons
- Depth can come as numeric too

## Load and massage the data

- Tidy up factors/levels (note the need to set new label on csv dataset)

- Separete in Sedling and regrowth phases

- Get data formats right


```{r LoadTidyUp, echo=FALSE,include=FALSE, echo=FALSE}

#dataRaw <- read.csv("C:\\GitHubRepos\\LP_analysis\\ObsPre.csv", header = TRUE, skip = 4)
dataRaw <- read.csv("K:\\CPDiary\\Data\\Lucerne & pasture drought trial 2011-2012\\ModelFitting\\ObsPre.csv", 
                    header = TRUE)
head(dataRaw)
summary(dataRaw)

# Create working df and sort out formats - 
dataWork <- as.data.frame(dataRaw)

# Force parameter values to factors
dataWork$Date <- as.Date(ymd(dataWork$Date))
dataWork$SurfaceKl <- as.factor(dataWork$SurfaceKl)
dataWork$RootFrontVelocity <- as.factor(dataWork$RootFrontVelocity)
dataWork$KLReduction <- as.factor(dataWork$KLReduction)

# sort depth out
dataWork$Depth <- gsub("D", "", dataWork$Depth)
dataWork$Depth <- factor(dataWork$Depth, 
                         levels = c(7, 23, 30, 50, 70, 90, 110, 130, 150))
dataWork$Depth <- as.factor(dataWork$Depth)

# Create separate Establishment and Regrowth periods
estDate <- "2012-06-01" # FIXME: find a more functional criteria later #estDate <- "2012-06-01 00:00:00"
dataWork$Stage <- as.factor(ifelse(dataWork$Date<=estDate,"Est","Reg"))



# Check levels
levels(dataWork$Depth)
levels(dataWork$SurfaceKl)
levels(dataWork$KLReduction)
levels(dataWork$RootFrontVelocity)

# Check
head(dataWork)
summary(dataWork)
str(dataWork)


# create indexes to call later
dataWork$index <- paste0(dataWork$Crop, "_",
                         dataWork$Stage, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index <- as.factor(dataWork$index)

# pooled by crop (excludes stage)
dataWork$index_crop <- paste0(dataWork$Crop, "_SK",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop <- as.factor(dataWork$index_crop)

# re-order fators (ensure Obs is the LAST one)
dataWork <- dataWork[c("Crop", "Irrigation",            
                       "Depth",
                       "Stage",
                       "Date",
                       "SurfaceKl", 
                       "KLReduction", 
                       "RootFrontVelocity", 
                       "index",
                       "index_crop",
                       "Pred", "Obs")]

```


## Check clean dataset


```{r, echo=FALSE}
# final product
str(dataWork)
head(dataWork, 20)
summary(dataWork)

```


## Compare Obs and simulated time-series 

- For "dryland"" crops just to check
- All simulated values (grey) and observed values (red)

```{r GraphExample, fig.height=18,fig.width=12, echo=FALSE}

# example of dryland crops
dataWork %>%
  filter(Irrigation =="None") %>%
  ggplot() +
  geom_point(aes(x=Date, y=Pred),colour="grey") +
  geom_point(aes(x=Date, y=Obs), size=2, colour="darkred") +
  facet_grid(Depth~Crop)

```

## XY graph for Prediced x Obs for whole dataset now

```{r XYgraph , fig.height=18,fig.width=12, echo=FALSE}
dataWork %>%
  ggplot(aes(x=Pred, y=Obs,colour=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
              linetype = 3,
              aes(colour=factor(Depth))) +
  geom_point() +
  facet_grid(Crop~Stage, scales ="free") +
  geom_abline(intercept = 0, slope = 1) 

```


```{r CreateStats, echo=FALSE, include=FALSE}

# Create stats function as per Gauch et al paper

gauchStats <- function(sim, meas) {
  
  n_s <- length(sim)
  n_m <- length(meas)
  model <- lm(meas~sim)
  sim_sq <- sum((sim - mean(sim))^2)
  mes_sq <- sum((meas - mean(meas))^2)
  r2 <- summary(model)$r.squared
  slope <- model$coefficients[[2]]
  
  sb <- (sum(mean(meas)) - sum(mean(sim)))^2
  nu <- (1-slope)^2 * (sim_sq/n_s)
  lc <- (1-r2) * (mes_sq/n_m)
  msd <- sb+nu+lc
  
  sb_r <- round((sb/msd)*100,1)
  nu_r <- round((nu/msd)*100,1)
  lc_r <- round((lc/msd)*100,1)
  
  msd_r <- sb_r+nu_r+lc_r
  
  out <- c(sb_r,nu_r,lc_r, msd_r)
 # out <- c(sb,nu,lc, msd,sim_sq,mes_sq,r2,n_s,n_m,slope ) # testing
  
  return(out)
  
}


# overall stats function including gauch FIXME: Not working
statsFunc <- function (sim, meas){
    n = n()
    r2 = br2(sim,meas)*100
    rmse = rmse(sim,meas)
    r_rmse = rmse(sim,Obs)/mean(meas)*100 # big precision to enable cretrial of selection 
    nse = NSE(sim,meas)
    sb = gauchStats(sim,meas)[1]
    nu = gauchStats(sim,meas)[2]
    lc = gauchStats(sim,meas)[3]
    
    out <- c(n, r2, rmse, r_rmse, nse, sb, nu, lc)
    
    return(out)
} 

dataWork %>%
  group_by(Crop, SurfaceKl, KLReduction, RootFrontVelocity, index, index_crop) %>%
  summarise(
    n = n(),
    r2 = br2(Pred,Obs)*100,
    rmse = rmse(Pred,Obs),
    r_rmse = rmse(Pred,Obs)/mean(Obs)*100, # big precision to enable cretrial of selection 
    nse = NSE(Pred,Obs),
    sb = gauchStats(Pred,Obs)[1],
    nu = gauchStats(Pred,Obs)[2],
    lc = gauchStats(Pred,Obs)[3]) 


# test dataset
s <- c(342.5,	68.3,	70.1,	286.1,	333.8)
m <- c(299.64,	161.36,	201.45,	220.8,	217.67)

gauchStats(s,m)

```

## Calculate table of statistics


```{r DoStatsAll, echo=FALSE, include=FALSE}

# Create a tailored common dplyr thread that do the stats for the df you have
doStats <-  function (x) {
    library(dplyr)
    x %>% dplyr::summarise(
    n = n(),
    r2 = br2(Pred,Obs)*100,
    rmse = rmse(Pred,Obs), # FIXME: what to do when there are similar rmse for different parameters?
    r_rmse = rmse(Pred,Obs)/mean(Obs)*100,
    nse = NSE(Pred,Obs),
    sb = gauchStats(Pred,Obs)[1],
  nu = gauchStats(Pred,Obs)[2],
  lc = gauchStats(Pred,Obs)[3])
}


# Pooled by Crop only - agregate and do stats across crops only
st1 <- dataWork %>%
  group_by(Crop, SurfaceKl, KLReduction, RootFrontVelocity, index_crop) %>%
  doStats()

summary(st1)

# Pooled by Crop & Stage - Agregate all layers (i.e. remove Depth from grouping) and do stats for Crop and Stage
st2 <- dataWork %>%
  dplyr::group_by(Crop, Stage, SurfaceKl, KLReduction, RootFrontVelocity, index) %>%
  doStats()
  
summary(st2)

# Stats per layer
# Test version - not used
# TODO: Do stats by soil depth and drop similations that have poor fit on top layers OR ...

st3 <- dataWork %>%
  dplyr::group_by(Crop, Stage, Depth, SurfaceKl, KLReduction, RootFrontVelocity, index) %>%
  doStats()
  
summary(st3)

# weights per layer
weightsPerLayer <- c(2,1,1,1,1,1,1,1,1)

# STOPPED here need to change values of st1 and st2 that where inverted
# Also need to get caegories numbers sorted

```

## Compare stats

```{r AnalyseStats, fig.height=18,fig.width=12, echo=FALSE}


# how are statistics related (exploratory stage)
# is there any relationship between r_rmse and lc
# st1 %>%
#   ggplot(aes(x=r_rmse , y=r2)) +
#   geom_point() + 
#   facet_grid(.~Crop, scales = "free")
  
# plots all stats against each other
plot(st1[,7:13])

st2  %>% 
  ungroup() %>% # to allow mutate
  # mutate(
  #   SurfaceKl = as.factor(SurfaceKl),
  #   KLReduction = as.factor(KLReduction),
  #   RootFrontVelocity = as.factor(RootFrontVelocity)
  # ) %>%
  group_by(Crop, Stage) %>%
  ggplot(aes(x = SurfaceKl, 
             y = KLReduction, 
             z = RootFrontVelocity, 
             fill = r_rmse)) +
  geom_tile() + 
  coord_equal() +
  stat_contour() + 
  facet_grid(Stage~Crop+RootFrontVelocity) + 
  scale_colour_gradientn(colours=terrain.colors(10)) + # FIXME: not working as wanted
  xlab("Surface kl category") +
  ylab("Exponential kl reduction category")+
  theme(text = element_text(size=16)) + theme(axis.text.x = element_text(angle = 90, hjust = -0.5))

# add continuous value sof factor (parameters)
# st1 <- merge(st1,kl_red_cats,by="KLReduction_cat")
# st2 <- merge(st1,kl_surface_cats,by="SurfaceKl")
# st2 <- merge(st1,RFV_cats,by="RootFrontVelocity")

# Sensitivity plots for lc and rmse
st2 %>%
  ungroup() %>% # to allow mutate
  # mutate(RootFrontVelocity = 
  #          factor(RootFrontVelocity,levels = c("Low", "Mid", "High"))) %>%
  ggplot(aes(x=r_rmse, colour=factor(SurfaceKl),fill=factor(SurfaceKl))) +
  geom_density(alpha=0.2) + 
  facet_grid(Stage~Crop)

head(st2)

# 



```


# Calculates the best fit (lowest r_rmse)


## For the "establishment phase only"


- Searches for best fit combination in establishment phase
- Selects root front velocity for that phase

```{r BestFitFinder1, echo=FALSE}

# Define a selection criteria

# NOT tested:
  # mutate(mse = rmse^2, nlc = 100-lc, mse_nlc = mse*nlc ) %>% # creates metrics for selecting best fit
  # filter(mse_nlc == min (mse_nlc)) %>%
  # filter(r_rmse == min(r_rmse) ) %>%
  # dplyr::select(-mse, -nlc, -mse_nlc)


# Define a quantile of rmse to explore populations
p <- 0.25

# st2 %>% group_by(Crop,Stage) %>% mutate(Q = quantile(rmse,probs=p)) %>% summary() # FIXME: does not calculate quartile per group

# workouround
Q_luc_est <- st2 %>% subset(Crop == "Lucerne" & Stage == "Est") %>% mutate(Q = quantile(rmse,probs=p))
Q_luc_reg <- st2 %>% subset(Crop == "Lucerne" & Stage == "Reg") %>% mutate(Q = quantile(rmse,probs=p))  
Q_pas_est <- st2 %>% subset(Crop == "Pasture" & Stage == "Est") %>% mutate(Q = quantile(rmse,probs=p)) 
Q_pas_reg <- st2 %>% subset(Crop == "Pasture" & Stage == "Reg") %>% mutate(Q = quantile(rmse,probs=p))  


# find quartiles for each crop/stage FIXME: Sure there's a way to do this in groups 
st2 <- st2 %>% mutate(rmseCut = ifelse(grepl("Lucerne_Est", index), Q_luc_est$Q[1],
                               ifelse(grepl("Lucerne_Reg", index), Q_luc_reg$Q[1],
                                   ifelse(grepl("Pasture_Est", index), Q_pas_est$Q[1],
                                      ifelse (grepl("Pasture_Reg", index), Q_pas_reg$Q[1], NA)))
                               ))
summary(st2)

# finds lowest r_rmse for the establishment phase first
bf0 <- st2  %>%
  filter(Stage == "Est") %>%
  group_by(Crop) %>%
  filter(rmse<=rmseCut) %>%
  filter(lc == max(lc)) %>%
  filter(r_rmse == min(r_rmse))

rownames(bf0) <- paste0(bf0$Crop)

t(bf0) %>%
  kable(format = "markdown")

selecParamsEst <- as.character(unique(bf0$index))
RVF_luc_est <- bf0[bf0$Crop=="Lucerne",]$RootFrontVelocity
RVF_rye_est <- bf0[bf0$Crop=="Pasture",]$RootFrontVelocity

# now select best fit for regrowth by setting the RFV to the selected in the establishment phase
bf1 <- st2  %>% 
  filter(  
    index == selecParamsEst[1] | # bring back only selected establishmnet combinations 
    index == selecParamsEst[2] |
    (Crop == "Lucerne" &  Stage == "Reg" & RootFrontVelocity == RVF_luc_est) |   
    (Crop == "Pasture" &  Stage == "Reg" & RootFrontVelocity == RVF_rye_est)) %>% 
  group_by(Crop, Stage) %>%
  filter(rmse<=rmseCut) %>%
  filter(lc == max(lc)) %>%
  filter(r_rmse == min(r_rmse) )
  
rownames(bf1) <- paste0(bf1$Crop,".",bf1$Stage)

t(bf1) %>%
  kable(format = "markdown")


```

## Best fit graphs for LUCERNE

```{r PlotBestFitsLuc, fig.height=18,fig.width=12, echo=FALSE}



# Input best fits here (FIXME: Automate this as a function from paramater estimation)

selecParams <- as.character(unique(bf1$index))

sel1 <- dataWork %>%
  filter(
    # establisment
    index == selecParams[1] | # watch for this conventions of names (FIXME)
    # regrowth phase
    index == selecParams[2]
  ) 

# finds the closest point to the date of establihment
estLoc <- which.min(abs(as.Date(sel1$Date) - as.Date(estDate)))

# graph lucerne
sel1 %>%
  ggplot() +
  geom_point(aes(x=Date, y=Obs), size=2, shape = 1) +
  geom_line(aes(x=Date, y=Pred),colour="black") +
  facet_grid(Depth~Irrigation, scales = "free") + 
  scale_x_date(labels = date_format("%b")) + scale_shape(solid = FALSE) + 
  geom_vline(xintercept = as.numeric(sel1$Date[estLoc]),linetype = 2) +
  ylab("Volumetric soil moisture (mm)") +
  theme(text = element_text(size=16))

#sel1  %>%
#  ggplot(aes(x=Pred, y=Obs,fill = factor(Stage),colour=factor(Stage))) +
#  stat_smooth(method = "lm", se = TRUE, 
#            linetype = 3) +
#  geom_point() +
#  facet_grid(Depth~Irrigation, scales = "free") +
#  geom_abline(intercept = 0, slope = 1) 

sel1  %>%
  ggplot(aes(x=Pred, y=Obs,fill = factor(Depth),colour=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
            linetype = 3) +
  geom_point() +
  facet_grid(Stage~Irrigation, scales = "free") +
  geom_abline(intercept = 0, slope = 1) 

```

## Best fit graphs for RYEGRASS

```{r BestFitRye, fig.height=18,fig.width=12, echo=FALSE}


# Ryegrass
sel2 <- dataWork %>%
  filter(
    # establisment
    index == selecParams[3] |
    # regrowth phase
    index == selecParams[4]
  ) 


# graph ryegrass
 sel2 %>%
  ggplot() +
  geom_point(aes(x=Date, y=Obs), size=2, shape = 1) +
  geom_line(aes(x=Date, y=Pred),colour="black") +
  facet_grid(Depth~Irrigation, scales = "free") + 
  scale_x_date(labels = date_format("%b")) + scale_shape(solid = FALSE) + 
  geom_vline(xintercept = as.numeric(sel1$Date[estLoc]),linetype = 2)+
  ylab("Volumetric soil moisture (mm)") +
  theme(text = element_text(size=16))

#sel2  %>%
#  ggplot(aes(x=Pred, y=Obs,fill = factor(Stage),colour=factor(Stage))) +
#  stat_smooth(method = "lm", se = TRUE, 
#            linetype = 3) +
#  geom_point() +
#  facet_grid(Depth~Irrigation, scales = "free") +
#  geom_abline(intercept = 0, slope = 1) 

sel2  %>%
  ggplot(aes(x=Pred, y=Obs,fill = factor(Depth),colour=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
            linetype = 3) +
  geom_point() +
  facet_grid(Stage~Irrigation, scales = "free") +
  geom_abline(intercept = 0, slope = 1)                  



```

## How does the best fit(s) look for RMSE by depth?

```{r RmsdPerDepth, echo = FALSE, fig.height=12,fig.width=12, echo=FALSE}

# calculate the stats for the BEST fit
# For each layer and depth combination to show trends
# IDEA: how much this afected water uptake?

sel3 <- rbind(sel1, sel2)

# calculates stats by depth
g1 <- sel3 %>%
  dplyr::group_by(Crop, Stage, Irrigation, Depth, Depth, SurfaceKl, KLReduction, RootFrontVelocity) %>%
  dplyr::summarise(
    n = n(),
    r2 = round(br2(Pred,Obs)*100,0),
    rmse = round(rmse(Pred,Obs),3),
    r_rmse = round(rmse(Pred,Obs)/mean(Obs)*100,3),
    nse = round(NSE(Pred,Obs),1),
    sb = gauchStats(Pred,Obs)[1],
  nu = gauchStats(Pred,Obs)[2],
  lc = gauchStats(Pred,Obs)[3]) 

head(g1,20)

# create a factor of  depth as integer
#g1$Depth <- as.numeric(gsub("D", "",g1$Depth)) # FIXME: move this to a higher level (earlier) as it can be used in other graphs

# _rmse by depth and crop
g1 %>%
  ggplot(aes(x=as.numeric(as.character(Depth))*-1, y=r_rmse, colour=factor(Stage))) +
  geom_point() +
  geom_line() +
 # geom_bar(aes(stats='identity')) +
  facet_grid(Crop~Irrigation) +
  coord_flip() + 
  geom_hline(yintercept = 20,linetype = 20) +
  labs(x = "Soil depth (cm)", y = "Relative Root Mean Squared Error (RMSE, % mean)") +
  theme(text = element_text(size=rel(4))) +
  theme(legend.text=element_text(size=16))

```


## Bring all stats together

- Compare stats across depths now

```{r StatsPerDepth, echo = FALSE, fig.height=12,fig.width=12, echo=FALSE}
# bring all stats together

# lucerne
g1 %>%
#  dplyr::select(-sb,-nu,-lc,-n) %>%
  tidyr::gather("StatsType","StatsValue", 8:ncol(g1)) %>%
  filter(Crop == "Lucerne") %>%
  ggplot(aes(x=as.numeric(as.character(Depth)), y=StatsValue, colour=factor(Stage))) +
  geom_point() +
  geom_line() +
 # geom_bar(aes(stats='identity')) +
  facet_wrap(~StatsType+Irrigation,scales = "free", ncol=4) +
 # coord_flip() + 
  labs(x = "Soil depth (cm)", y = "Stats (xxxxx)") +
  theme(text = element_text(size=rel(4))) +
  theme(legend.text=element_text(size=16))


# pasture
g1 %>%
 # dplyr::select(-sb,-nu,-lc,-n) %>%
  tidyr::gather("StatsType","StatsValue", 8:ncol(g1)) %>%
  filter(Crop == "Pasture") %>%
  ggplot(aes(x=as.numeric(as.character(Depth)), y=StatsValue, colour=factor(Stage))) +
  geom_point() +
  geom_line() +
  facet_wrap(~StatsType~Irrigation, scales = "free") +
#  coord_flip() + 
  labs(x = "Soil depth (cm)", y = "Stats (xxxxxx)") +
  theme(text = element_text(size=rel(4))) +
  theme(legend.text=element_text(size=16))
```

## Partition of RMSE into Gauch fractions

- Standard bias (SB)

- Non Unit (NU)

- Lack of Correlation (LU)

```{r GauchStatsByDepth, echo = TRUE, fig.height=12,fig.width=12, echo=FALSE}

df <- g1 %>%
  mutate(mse = rmse^2, sb=sb*mse, nu=nu*mse, lc=lc*mse) %>%
 # dplyr::select(-n, -r2,-rmse,-r_rmse,-nse) %>%
  dplyr::select(-n, -r2,-r_rmse,-nse, -rmse) %>%
  tidyr::gather("StatsType","StatsValue", 8:10) # %>%
  # mutate(Depth = factor(Depth, rev(cat_depth)))

# graph
  df %>%
#  filter(Crop == "Lucerne") %>%
  ungroup() %>%  
  mutate(Depth = factor(Depth, level = rev(Depth))) %>%
  ggplot(aes(x=Depth, y=StatsValue,fill=rev(StatsType))) +
  geom_bar(stat='identity') +
  facet_grid(Stage+Crop~Irrigation,scales = "free") +
  coord_flip() + 
  labs(x = "Soil depth (cm)", y = "MSE") +
  theme(text = element_text(size=rel(4))) +
  theme(legend.text=element_text(size=16))

```

```{r PinPointBestFit}

dfTemp <- st2 %>%
  group_by(Crop, Stage) %>%
  dplyr::select(lc, r_rmse) %>%
  summarise_each(funs(max, min)) %>%
  dplyr::select(Crop, Stage, lc_max,r_rmse_min)

st2 %>%
  mutate(BestFit = ifelse(index %in% bf1$index,"Best", "NotBest")) %>%
  mutate(BestFit = factor(BestFit, levels=c("NotBest", "Best"))) %>%
  ggplot(aes(x=lc, y=r_rmse)) +
  geom_point(aes(colour=factor(BestFit), size=factor(BestFit), shape=factor(BestFit))) +
  facet_grid(Stage~Crop, scales= "free") +
  xlab("Lack of correlation (LC, %MSE) ") +
  ylab(" Relative RMSE (% mean)") +
  geom_hline(data=dfTemp, aes(yintercept=r_rmse_min), linetype=2) +
  geom_vline(data=dfTemp, aes(xintercept=lc_max), linetype=2)


```

