---
title: "Analysis of L&P dataset"
output:
  html_document: default
  word_document: default
---

```{r LoadLibs, echo=FALSE, include=FALSE}
library(xtable)
library(knitr)
library(RSQLite)
library(lubridate)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library (hydroGOF)
library(scales)
library(GGally)
library(xtable)

options("scipen"=100, "digits"=4) # prints not in scientific notation

```

## FIXME list

- Print numeric value of Surfacekl in csv output
- Use numeric value of parameters as characters througout to simplify comparisons
- Depth can come as numeric too
- The current criteria for parameter selection has HUGE leverage on results
- How to define best fit?????

## Load and massage the data

- Tidy up factors/levels (note the need to set new label on csv dataset)

- Separete in Sedling and regrowth phases

- Get data formats right


```{r ReadData, echo=FALSE,include=TRUE, echo=FALSE}

#dataRaw <- read.csv("C:\\GitHubRepos\\LP_analysis\\ObsPre.csv", header = TRUE, skip = 4)
# dataRaw <- read.csv("K:\\CPDiary\\Data\\Lucerne & pasture drought trial 2011-2012\\ModelFitting\\ObsPre.csv", 
#                     header = TRUE)

dataRaw <- read.csv("K:\\CPDiary\\Data\\Lucerne & pasture drought trial 2011-2012\\ModelFitting\\ObsPreSD15.csv", 
                    header = TRUE)
#head(dataRaw)
summary(dataRaw)
```


# merge with observation per replicate


```{r}
temp_obs <- read.csv("K:\\CPDiary\\Data\\Lucerne & pasture drought trial 2011-2012\\ModelFitting\\worked_observed_data.csv")

head(temp_obs)
dataRaw_merge <- merge(dataRaw,temp_obs, by=c("Depth","Crop","Irrigation", "Date"))
  
```
```{r}
head(dataRaw_merge)
```


```{r dataTidy}

# Create working df and sort out formats - 
dataWork <- as.data.frame(dataRaw_merge)

# Force parameter values to factors
dataWork$Date <- as.Date(ymd(dataWork$Date))
dataWork$SurfaceKl <- as.factor(dataWork$SurfaceKl)
dataWork$RootFrontVelocity <- as.factor(dataWork$RootFrontVelocity)
dataWork$KLReduction <- as.factor(dataWork$KLReduction)

# sort depth out
dataWork$Depth <- gsub("D", "", dataWork$Depth)
dataWork$Depth <- factor(dataWork$Depth, 
                         levels = c(7, 23, 30, 50, 70, 90, 110, 130, 150))
dataWork$Depth <- as.factor(dataWork$Depth)

# Create separate Establishment and Regrowth periods
estDate <- "2012-06-01" # FIXME: find a more functional criteria later




#estDate <- "2012-01-01" # FIXME: test sensitivity to this date
#estDate <- "2012-06-01 00:00:00"
dataWork$Stage <- as.factor(ifelse(dataWork$Date<=estDate,"Establishment","Regrowth"))
# sort out irrigation treat names
dataWork <- dataWork %>% mutate(Irrigation = factor(Irrigation, 
                                                    levels=c("None", 
                                                             "ThreeWeekly",
                                                             "OnePerWeek",
                                                             "TwoPerWeek"))) # re-order factors
levels(dataWork$Irrigation) <- c("No Irrigation",
                                 "Low Frequency",
                                 "Medium Frequency", 
                                 "High Frequency") # rename factors, attention to match!
# sort out crop names
dataWork <- dataWork %>% mutate(Crop=factor(Crop, levels=c("Lucerne", "Pasture"))) # re-order factors
levels(dataWork$Crop) <- c("Lucerne","Ryegrass") # rename factors, attention to match!

# revove unrealistic levels (FIXME: decibe about this after discussing with Hamish)
#dataWork <- dataWork %>% filter(SurfaceKl != "20") 

# Check levels
levels(dataWork$Depth)
levels(dataWork$SurfaceKl)
levels(dataWork$KLReduction)
levels(dataWork$RootFrontVelocity)

```


```{r}
# Check
#head(dataWork)
summary(dataWork)

```

```{r}
str(dataWork)
```


## Check clean dataset


```{r, echo=FALSE}
# final product
str(dataWork)

```

```{r, include = TRUE}
head(dataWork, 20)
```

```{r}
summary(dataWork)
```

## Check Observed and Simulated time-series

- Show time-series

- Simulated parameter combinations' data (grey) and observed values (red)

- Note that, in some cases, observations are at the edge of simulation's range. 

- <span style="color:red">To discuss: Is this a sign that parameter ranges should be expanded???</span>  

```{r GraphExample, fig.height=8,fig.width=15, echo=FALSE}

# example of dryland crops
dataWork %>%
 # filter(Irrigation == "No Irrigation") %>%
  ggplot() +
  geom_point(aes(x=Date, y=Pred),colour="grey") +
  geom_point(aes(x=Date, y=Obs), size=1.5, colour="darkred") +
  facet_grid(Depth~Crop+Irrigation) +
  geom_vline(xintercept = as.numeric(as.Date(ymd(estDate))), linetype=2) 

```

```{r RepVariab, include=FALSE, echo=FALSE}

# FIXME: Using mock values of observed values per rep until get atual from Hamish

# This will be deleted one we have actual values available

# df_mock <- dataWork %>%
#   filter(SurfaceKl == "3", KLReduction=="-0.01", RootFrontVelocity=="5") %>% # get a single parameter combination
#   mutate(Rep_1 = Obs*runif(1, min = 0.8, max = 1.2)) %>%
#   mutate(Rep_2 = Obs*runif(1, min = 0.8, max = 1.2)) %>%
#   mutate(Rep_3 = Obs*runif(1, min = 0.8, max = 1.2)) %>%
#   mutate(Rep_4 = Obs*runif(1, min = 0.8, max = 1.2)) %>%
#   rowwise() %>%
#   mutate(Obs_av = sum(Rep_1,Rep_2,Rep_3,Rep_4)/4)  %>%
#   mutate(CI95_down = (Obs_av-3.182*sd(c(Rep_1,Rep_2,Rep_3, Rep_4)))) %>% # FIXME: check with Esther how to
#   mutate(CI95_up = (Obs_av+3.182*sd(c(Rep_1,Rep_2,Rep_3, Rep_4)))) # calculate 95% CI properly
# 
# # head(df_mock)

```

## Plot observations and uncertanty
```{r, include=FALSE, echo=FALSE}
# check if moch average is close to real average
# plot(df_mock$Obs_av,df_mock$Obs)
```

```{r, fig.height=15,fig.width=15, echo=FALSE}
# 
# df_mock %>%
#   dplyr::select(-Stage,-(SurfaceKl:Obs)) %>%
#   dplyr::select(Crop, Date, Irrigation,Depth, Rep_1:Obs_av) %>%
#   tidyr::gather("Rep","Obs",Rep_1:Obs_av) %>% # FIXME: check names to be used in raw data
#   mutate(Rep = factor(Rep)) %>%
#   filter(Rep != "Obs_av") %>% 
#   ggplot(aes(x=Date, y=Obs)) +
#   geom_point(aes(colour=factor(Rep)), alpha=0.5) +
#   geom_smooth(colour="black", size=1.1) +
#   geom_line(data=df_mock, aes(x=Date, y=CI95_down), linetype=2, colour="darkgrey") +
#   geom_line(data=df_mock, aes(x=Date, y=CI95_up), linetype=2, colour="darkgrey") +
#   facet_grid(Depth~Crop+Irrigation) + 
# #  scale_x_date(labels = date_format("%b")) + scale_shape(solid = FALSE) + 
#   ylab("Volumetric soil moisture (mm)") +
#   theme(text = element_text(size=18)) +
#     labs(colour="Experiental replicate") + theme(legend.position="top")
```

## Graph Predicted x Observed

- For whole dataset (all parameter combinations)
- 1:1  line in black (x=y)
- linear regression for each soil depth shown in gray
- 95% confidence interval also shown

```{r XYgraph , fig.height=10,fig.width=12, echo=FALSE, include = TRUE}
dataWork %>%
  ggplot(aes(x=Pred, y=Obs,colour=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
              linetype = 3,
              aes(colour=factor(Depth))) +
  geom_point(alpha = 0.2) +
  facet_grid(Crop~Stage, scales ="free") +
  geom_abline(intercept = 0, slope = 1) + 
  coord_fixed(ratio = 1) +
  xlim(0,max(dataWork$Pred,dataWork$Obs)) +
  ylim(0,max(dataWork$Pred,dataWork$Obs))

```

## Create Gauch stats function

- Function to calculate RMSE components
- Standard bias (SB)
- Non unit (NU)
- Lack of correlation (LC)
- Reference: Gauch Jr, H.G., Hwang, J.T.G., Fick, G.W., 2003. Model evaluation by comparison of model-based predictions and measured values. Agron. J. 95, 1442-1446

```{r CreateStats, echo=FALSE, include=FALSE}

# Create stats function as per Gauch et al paper

gauchStats <- function(sim, meas) {

  n_s <- length(sim)
  n_m <- length(meas)
  model <- lm(meas~sim)
  sim_sq <- sum((sim - mean(sim))^2)
  mes_sq <- sum((meas - mean(meas))^2)
  r2 <- summary(model)$r.squared
  slope <- model$coefficients[[2]]

  sb <- (sum(mean(meas)) - sum(mean(sim)))^2
  nu <- (1-slope)^2 * (sim_sq/n_s)
  lc <- (1-r2) * (mes_sq/n_m)
  msd <- sb+nu+lc

  sb_r <- round((sb/msd)*100,1)
  nu_r <- round((nu/msd)*100,1)
  lc_r <- round((lc/msd)*100,1)

  msd_r <- sb_r+nu_r+lc_r

  # select which variables to output
  out <- c(sb_r, # output 1
           nu_r,
           lc_r, 
           msd_r, 
           round(r2*100,1)) # output 5

  return(out)

}

```

- Test Gauch stats
- Result has to be: 76.5  5.2 18.2 99.9  7.1
- Result is:

```{r, include=TRUE, echo=FALSE}
# test dataset
s <- c(4231.972,3935.604,3779.652,3627.687,3363.499,3230.566,2868.114,2868.827)
m <- c(4987.66,5636.09,4754.06,4114.53,4141.72,3704.06,5142.19,4762.03)

gauchStats(s,m)
```

## Creates a generic set of stats to output 

- These are the standard stats used in paper
- Basic stats (e.g. r2, RMSE ...) plus the Gauch et al. components defined in the function above

```{r, echo = TRUE, include = FALSE}

# Create a tailored common function that do the stats and includes the gauch functionality
# input is a df with Pred and Obs columns

doStats <-  function (x) {
    library(dplyr)
    x %>% dplyr::summarise(
    n = n(),
    r2 = gauchStats(Pred,Obs)[5],
    rmse = rmse(Pred,Obs),
    r_rmse = rmse(Pred,Obs)/mean(Obs)*100, # big precision to enable cretrial of selection
    nse = NSE(Pred,Obs),
    sb = gauchStats(Pred,Obs)[1],
    nu = gauchStats(Pred,Obs)[2],
    lc = gauchStats(Pred,Obs)[3],
    rsr = rsr(Pred,Obs), # Ratio of RMSE to the standard deviation of the observations
    pBias = pbias(Pred,Obs)) # Percent Bias between sim and obs, with treatment of missing values.
}

```

## Create indexes necessary for data aggregation to do stats

- Aggregate Pred x Obs data by different "factors" to apply stats
- For example, to test best parameter combinations for "each" crop and "each" Stage, use: - i.e. calculates stats across "all" soil depths and "all" water-treatments for each "Crop and Stage" combination
- Multiple agregations are tested below

```{r CreateIndexes, echo=TRUE}

# "Crop and Stage" calculates stats across "all" soil depths and "all" water-treatments 
# for each "Crop and Stage" combination

# Best fit by crop only ---------------------------------------
dataWork$index_crop <- paste0(dataWork$Crop, "_SK",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop <- as.factor(dataWork$index_crop)


# Best fit  by crop and "stage" ---------------------------------------------
dataWork$index_crop_stage <- paste0(dataWork$Crop, "_",
                         dataWork$Stage, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_stage <- as.factor(dataWork$index_crop_stage)

# Best fit  by crop and "depth" ---------------------------------------
dataWork$index_crop_depth <- paste0(dataWork$Crop, "_",
                              dataWork$Depth, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_depth <- as.factor(dataWork$index_crop_depth)

# Best fit  by crop and irrigation ---------------------------------------
dataWork$index_crop_irrig <- paste0(dataWork$Crop, "_",
                              dataWork$Irrigation, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_irrig <- as.factor(dataWork$index_crop_irrig)

# Best fit  by crop, stage and depth ---------------------------------------
dataWork$index_crop_stage_depth <- paste0(dataWork$Crop, "_",
                              dataWork$Stage, "_Dp:",
                              dataWork$Depth, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_stage_depth <- as.factor(dataWork$index_crop_stage_depth)


# Best fit  by crop and stage and irrigation ---------------------------------------
dataWork$index_crop_stage_irrig <- paste0(dataWork$Crop, "_",
                              dataWork$Stage, "_Irr:",
                              dataWork$Irrigation, "_SK:",
                              dataWork$SurfaceKl, "_KR:",
                              dataWork$KLReduction, "_RF:",
                              dataWork$RootFrontVelocity)

dataWork$index_crop_stage_irrig <- as.factor(dataWork$index_crop_stage_irrig)


# re-order fators for tidy up (ensure Obs is the LAST one)
dataWork <- dataWork[c("Crop", "Irrigation",            
                       "Depth",
                       "Stage",
                       "Date",
                       "SurfaceKl", 
                       "KLReduction", 
                       "RootFrontVelocity",
                       "index_crop",
                       "index_crop_stage",
                       "index_crop_depth",
                       "index_crop_irrig",
                       "index_crop_stage_depth",
                       "index_crop_stage_irrig",
                       "Pred", "Obs","CI95_down","CI95_up")]

# save it for the sensitivity analysis
write.csv(dataWork, "C:\\GitHubRepos\\FodderBeetRUE_2016\\workedData.csv")
```

## Calculate statistics

# Percent of prediction within the 95% CI

- this stats has to be calculated separatelly and later merged with reported stats

```{r, warning = FALSE, echo = FALSE}


tempDF <- dataWork %>%
 # mutate(CI95_down = 0.95*Obs, CI95_up = 1.05*Obs) %>% # FIXME: Test CI calc
  rowwise() %>%
  mutate(Pred_gt_CId = ifelse(Pred>CI95_down,1,0)) %>% 
  mutate(Pred_lw_CIu = ifelse(Pred>CI95_down,1,0), 
         Within_CI95 = Pred_gt_CId*Pred_lw_CIu) %>%
  dplyr::select(-Pred_gt_CId,-Pred_lw_CIu)

```

- Graph shows distribution of % of points within the 95% CI of observed


```{r, fig.height=10, fig.width=12, warning = FALSE, include = TRUE}

# This has to be done for stats batch "best" fit or stats calculation
tempDF  %>%
group_by(Crop, Stage,Irrigation,RootFrontVelocity,KLReduction,SurfaceKl, index_crop_stage_irrig) %>%
summarise(Perc95CI = sum(Within_CI95)/n()*100, ObsNo = n()) %>%
  ggplot(aes(x=factor(SurfaceKl) , y=Perc95CI)) +
  geom_boxplot(alpha=0.2) +
  geom_jitter(width=0.5, alpha=0.5, aes(shape=factor(RootFrontVelocity), colour=factor(KLReduction))) +
  facet_grid(Irrigation~Crop+Stage) 

```

## Calculate stats across data subsets

Selection of stat indexes to use, at the moment we're following the refs:

(i) Moriasi, D.N., Arnold, J.G., Van Liew, M.W., Bingner, R.L., R.D., H., Veith, T.L., 2007. Model evaluation guidelines for systematic quantification of accuracy in watershed simulations. Transactions of the ASABE 50, 885-900

(ii) Gauch Jr, H.G., Hwang, J.T.G., Fick, G.W., 2003. Model evaluation by comparison of model-based predictions and measured values. Agron. J. 95, 1442-1446

Logic used:

- First calculates stats for each "crop by stage by irrigation" combination

- Will get RFV for dryland from here
 
```{r DoStatsCrop, echo=FALSE, include=TRUE}

# IDEA: compare optimum for all x optimuns for stage/irrig separation to test assumption

# Separate Crop only
statDF_crop_stage_irrig <- dataWork  %>%
#  filter(Irrigation == "No Irrigation") %>%  # optmise for dryland only
# explore param diffs for stage and irrig treat (test only)  
# filter(Irrigation != "No Irrigation") %>% 
# filter(Irrigation != "High frequency") %>% 
# filter (Stage == "Establishment") %>%
# filter (Stage == "Regrowth") %>%
group_by(Crop, Stage, Irrigation, # labels
         index_crop, # used later
         index_crop_stage_irrig, # corresponding index
         SurfaceKl, KLReduction, RootFrontVelocity) %>% # index
doStats() 

summary(statDF_crop_stage_irrig)
```

# Define stats to be used for "optimum" parameter selection

- Use the stats calculated above

- <span style="color:red">Which method do we use to calibrate/optimise parameters?</span>

- Although there are sure many "standard" options: See Chapter 6 in Wallach, D., Makowski, D., Jones, J.W., 2013. Working with Dynamic Crop Models : Methods, Tools and Examples for Agriculture and Environment (2). Academic Press, Amsterdam, NL. ...

- ... we still need to "decide" what a "best-fit" is. 

- Many of the indexes are "highly" correlated - i.e. they are measures of same aspects (e.g. nse x r_rmse x rmse)

- So we propose to combine stats that have low correlation with each other, as they capture different aspects 

- Colours are the diffrenet crops lucerne (red) and ryegrass (blue)

```{r CompareStatIndexes, fig.height=8,fig.width=12, echo=FALSE, warning=FALSE, }
# with stats by Crop
statDF_crop_stage_irrig %>%
group_by(Crop) %>%
dplyr::select(Crop,r2:pBias) %>%
ggpairs(aes(colour=Crop),labelHJust=1)
```

For our situation, where we started from pre-simulated values for parameter combinations.

the proposed logic for parameter selection is:

- Two indexes that capture (i) simulation error and (ii) simulation bias

- We choose two that have weak correlation between themselves:(i) Option 1 (RMSE + LC) from Gauch et al 2003 or (ii) Option 2 (RSR + pBias) from Moriase et al 2007

-<span style="color:red">Is this appropriate? Any other ideas?</span>

# Current approach:

- <span style="color:red">Isolate the upper quartile of lack of correlation (LC)</span>: lowest Bias + Non-Unit

- <span style="color:red">select lowest rRMSE value</span>: lowest error

- <span style="color:red">get lowest parameter value there's if more than one "best-fit"</span>: pragmatism

- <span style="color:red">Is this a good rationale? Alternatives?</span>

# In practice:

- Subset the 25% "less biased" simulations (i.e. low SB or pBias close to zero)

- From these, find the one with lowest error (low RMSE or RSR)

- If multiple simulations fulfill criteria, take the one with lowest numeric parameter values (FIXME)

The choice of any best fit if multiple have same stats is due to equifinality - that will be explicitly dealt with later.

Table below show the best fit (parameter-combination) for each crop/stage/irrigation


```{r SelectBestSims, warning=FALSE}

# get stats for all combinations and filter for dry when needed

# idea ... loop across stages and treatment combinations
# idea: use percentile to get parameter uncertanty 
# use histogram of estimated values?
# FIXME: It is unclear how to combine stats criteria to select what the best simulation is

# FIXME: change names what's the diff between bestFitsims and best_fit_all?

percSel <- 0.25 # percentile of simulations to be selected (Attention to index criteria)

# finds lowest r_rmse for the establishment phase first
bestFit_sims <- statDF_crop_stage_irrig  %>%
  group_by(Crop, Irrigation, Stage) %>%
 # mutate(StatIndex = (100-lc)*r_rmse) %>%
 # mutate(StatIndex = pBias * rsr)  %>%
 # filter(StatIndex == min(StatIndex)) %>%
# filter(lc > quantile(lc, (1-percSel)) ) %>% # get least biased fits
 filter(r_rmse < quantile(r_rmse, (percSel)) ) %>% #
#  filter(pBias < ifelse(pBias<0, quantile(-1*pBias, 0.95), quantile(pBias, 0.95))) %>% # get 25% least biased fits
#  filter(r_rmse == min(r_rmse) ) %>%   # get smallest error
# filter(r_rmse == min(r_rmse) ) %>%   # get smallest error
 filter(lc == max(lc) ) %>%   # get sm
#  filter(lc > quantile(lc, (1-percSel)) | r_rmse < quantile(rmse, percSel))
  # mutate(pBias_pos = ifelse(pBias<0, -1*pBias, pBias)) %>% # adjust pBias to be always positive
  # filter(pBias_pos < quantile(pBias_pos, 0.05) | rsr < quantile(rsr, 0.05))
#  filter(rsr == min(rsr) ) %>%   # get smallest error
  filter(row_number() == min(row_number()))# FIXME: gets lowest param value if stats are the similar for many combinations 

#write.csv(bestFit_sims, "testStatsBias.csv")

# only for single selection
 rownames(bestFit_sims) <- paste0(substr(bestFit_sims$Crop,1,3),"_",
                                  substr(bestFit_sims$Stage,1,3),"_",
                                  substr(bestFit_sims$Irrigation,1,3))

# re-arrange as table df
 df <-  as.data.frame(t(bestFit_sims)) %>%
   mutate(Item = rownames(.))
  
 df %>%
    filter(Item != "index_crop_stage_irrig") %>%
    kable(format = "markdown", align = 'c')
  # xtable()
 

 # Isolate all best fits (FIXME: is this still used?)
 selecParamsEst <- as.character(unique(bestFit_sims$index_crop_stage_irrig))

```


```{r, fig.height=10, fig.width=15, warning = FALSE, include=FALSE, echo = FALSE}

# Evaluate residuals of best fit selection (stored for re-use if needed)
# 
# - Visually it seems "rainfed" simulations are less biased
# - This will be quantitativelly picked by bias indexes I believe (Check with Esther) 

dataWork %>%
  filter(index_crop_stage_irrig %in% as.character(bestFit_sims$index_crop_stage_irrig)) %>%
#  mutate(Res = (Pred-Obs)/Obs*100) %>%
  mutate(Res = Pred-Obs) %>%
  ggplot(aes(x=Obs,y=Res, colour=factor(Irrigation))) +
  geom_point(shape=21, alpha=0.8) +
  geom_smooth(method='lm', alpha=0.5) +
  geom_hline(yintercept=0, linetype=2, colour='darkgrey') +
  facet_grid(Depth~Crop+Stage) +
  ylab("Residue (mm/mm)") +
  xlab("Observed mean value (mm/mm) ")
```

## Select the best-fit

Select best fit for:

(i) Dryland crops (rationale is that "maximum" kl is expressed only under dry conditions)


(ii) We select "best" Root Front Velocity (RFV) from the "establishment" stage only (rationale is that simulations are only sensitive to the parameters during this stage ... when roots actually grow)

# Repeat the stats calculation for crop by irrigation (i.e. across stages)

- Fix the RFV value found for the "establishment" period in dryland crops

```{r}

 # Isolate the best RFV value from dryland and establishmnet phase
 RVF_luc_est <- bestFit_sims[bestFit_sims$Crop=="Lucerne" & 
                               bestFit_sims$Irrigation =="No Irrigation" &
                               bestFit_sims$Stage =="Establishment",]$RootFrontVelocity
 
 RVF_rye_est <- bestFit_sims[bestFit_sims$Crop=="Ryegrass" & 
                               bestFit_sims$Irrigation =="No Irrigation" &
                               bestFit_sims$Stage =="Establishment",]$RootFrontVelocity


# calculate crop by irrigation stats (across stages)
statDF_crop_irrig <- dataWork  %>%
filter(
    (Crop == "Lucerne" & RootFrontVelocity == RVF_luc_est) |   
    (Crop == "Ryegrass" & RootFrontVelocity == RVF_rye_est)) %>% # "optimum" RFV for "establishmnent" 
group_by(Crop, Irrigation, # labels
         index_crop,       # Index at crop level
         index_crop_irrig, # corresponding index
         SurfaceKl, KLReduction, RootFrontVelocity) %>% # index
doStats() 

summary(statDF_crop_irrig)
```

# Select best parameters for dryland conditions (fixed RFV)

```{r}

percSel <- 0.25 # percentile of simulations to be selected (Attention to index criteria)

bestFit_dry <- statDF_crop_irrig  %>%
 group_by(Crop, Irrigation) %>%
 filter(r_rmse < quantile(r_rmse, (percSel))) %>%
 filter(lc == max(lc)) %>%   # get sm
 filter(row_number() == min(row_number())) %>% # FIXME: gets lowest param value for similar
 filter(Irrigation == "No Irrigation") # keeps only the dry treats

summary(bestFit_dry)

```

# show best fit dry

```{r, echo = FALSE}
# re-arrange as table df
 df <-  as.data.frame(t(bestFit_dry)) %>%
   mutate(Item = rownames(.))
  
 df %>%
    filter(Item != "index_crop_stage_irrig") %>%
    kable(format = "markdown", align = 'c')
  # xtable()
 
 write.csv(df, "BestFirDryland.csv")
```

# Graph best fits

Isolate the simulations for best parameters

```{r, echo = FALSE}

# isolate all data from "best" simulations
simsWithBestFitForDry  <- dataWork %>%
  filter(index_crop %in% as.character(bestFit_dry$index_crop)) 

# finds the closest point to the date of establihment
estLoc <- which.min(abs(as.Date(simsWithBestFitForDry$Date) - as.Date(estDate)))

summary(simsWithBestFitForDry )

```

# Graph lucerne

```{r, fig.width=12, fig.height=12, echo = FALSE}

# Graph lucerne
simsWithBestFitForDry  %>%
#  filter(Crop == thisCrop) %>%
  ggplot() +
#  geom_point(data=dataWork[dataWork$Crop == thisCrop,], aes(x=Date, y=Pred), colour="darkgrey",alpha=0.1) + # add shade of all sims
  geom_point(data=dataWork,aes(x=Date, y=Pred), colour="darkgrey",alpha=0.1) +
  geom_point(aes(x=Date, y=Obs), size=2, shape = 1, colour="darkred") +
  geom_line(aes(x=Date, y=Pred),colour="black", size=1.2) +
#  geom_point(aes(x=Date, y=Pred),colour="black", size=0.2) +
 # facet_grid(Depth~Irrigation, scales = "free") + 
  facet_grid(Depth~Crop+Irrigation) + 
  scale_x_date(labels = date_format("%b")) + scale_shape(solid = FALSE) + 
  geom_vline(xintercept = as.numeric(simsWithBestFitForDry $Date[estLoc]),linetype = 2, colour="darkgrey") +
  ylab("Volumetric soil moisture (mm)") +
  theme(text = element_text(size=13))
```

## XY graph of above best fit dataset 

```{r, fig.width=10, fig.height=8, echo = FALSE}

simsWithBestFitForDry %>%
  ggplot(aes(x=Pred, y=Obs, group = factor(Depth),fill=factor(Depth))) +
  stat_smooth(method = "lm", se = TRUE, 
              linetype = 3,aes(colour=factor(Depth))) +
  geom_point(aes(colour=factor(Depth)), alpha=0.5) +
  facet_grid(Irrigation~Crop+Stage, scales = "free") +
  geom_abline(intercept = 0, slope = 1) + 
  coord_fixed(ratio = 1) +
  xlim(0,max(simsWithBestFitForDry$Pred,simsWithBestFitForDry$Obs)) +
  ylim(0,max(simsWithBestFitForDry$Pred,simsWithBestFitForDry$Obs)) +
  theme(text = element_text(size=13))+
  ylab("Observed volumetric soil water (mm/mm)") +
  xlab("Predicted volumetric soil water (mm/mm)") +
  labs(fill="Soil depth (cm)") + guides(colour=FALSE) 
```


## How are stats for all irrigation treatments 

- Using best-fit for dryland
- Using RFV for dryland establishment

```{r, echo=FALSE}

# calculate crop by irrigation stats (across stages)
statDF_simsWithBestFitForDry  <- simsWithBestFitForDry  %>%
group_by(Crop, Stage, Irrigation, # labels
         index_crop,       # Index at crop level
         index_crop_stage_irrig, # corresponding index
         SurfaceKl, KLReduction, RootFrontVelocity) %>% # index
doStats() 

#summary(statDF_simsWithBestFitForDry )

# statDF_simsWithBestFitForDry  %>%
#   tidyr::gather("StatName", "StatValue",rsr:pBias) %>%
#   mutate(StatName = factor(StatName)) %>%
#   ggplot(aes(x=Irrigation, y=StatValue))+
#   geom_bar(stat='identity') +
#   facet_wrap(Crop+Stage~StatName, scales = "free",ncol = 4)



```

# print table of stats across all tretaments

```{r, echo = FALSE}

tableOfStats <- statDF_simsWithBestFitForDry  %>%
  as.data.frame() %>%
  dplyr::select(-(index_crop:RootFrontVelocity)) 

tableOfStats %>%
  t() %>%
  write.csv("TableStats.csv")

tableOfStats %>%
#  t() %>%
#  as.data.frame() %>%
  kable(digits=3)


```

## How are statistics across different "soil depths" 

```{r }
# calculate Gauch stats with BEST fit dry simulations
# For each layer and depth combination

# calculates stats by depth
gauch_depth <- simsWithBestFitForDry  %>%
  dplyr::group_by(Crop, Stage, Irrigation, Depth, Depth, SurfaceKl, KLReduction, RootFrontVelocity) %>%
  dplyr::summarise(
    n = n(),
    r2 = gauchStats(Pred,Obs)[5],
    rmse = round(rmse(Pred,Obs),3),
    r_rmse = round(rmse(Pred,Obs)/mean(Obs)*100,3),
    nse = round(NSE(Pred,Obs),1),
    sb = gauchStats(Pred,Obs)[1],
  nu = gauchStats(Pred,Obs)[2],
  lc = gauchStats(Pred,Obs)[3]) 

summary(gauch_depth)
```

# summarise stats

- Ranges of SB found

```{r}
gauch_depth %>%
  filter(Depth=="30") %>%
  dplyr::select(Crop, Stage, Irrigation, sb,nu,lc) %>%
  group_by(Crop,Irrigation) %>%
  dplyr::select(-(Depth:KLReduction), -(Stage), -(nu:lc))  %>%
  summarise_each(funs(min,max))
  
```

# How are the stats by treatment and depth?

```{r, fig.width=15, fig.height=15, warning=FALSE, echo = FALSE }
df <- gauch_depth %>%
  mutate(mse = rmse^2, sb=sb*mse, nu=nu*mse, lc=lc*mse) %>%
 # dplyr::select(-n, -r2,-rmse,-r_rmse,-nse) %>%
  dplyr::select(-n, -r2,-r_rmse,-nse, -rmse) %>%
  tidyr::gather("StatsType","StatsValue", sb:mse) %>%
  mutate(StatsType = factor(StatsType, levels = c("sb","nu","lc", "mse")))

# graph
  df %>%
#  filter(Crop == "Lucerne") %>%
  ungroup() %>%  
  mutate(Depth = factor(Depth, level = rev(Depth))) %>%
  filter(StatsType != "mse") %>%
  ggplot(aes(x=Depth, y=StatsValue,fill=StatsType)) +
  geom_bar(stat='identity') +
  facet_grid(Irrigation~Crop+Stage,scales = "free") +
  coord_flip() + 
  labs(x = "Soil depth (cm)", y = "Mean square error (MSE)") +
#  theme(text = element_text(size=rel(4))) +
#  theme(legend.text=element_text(size=16)) + 
    theme(legend.title=element_blank()) +
    theme(text = element_text(size=20)) +
    theme(legend.position="top")
  
```

## Show population of simulations

- Equifinality aspects to analyse
- i.e. Our "best-fit" is only one of many "good-enough" parameter combinations

Flag "best-fit" parameters across simulations for each "crop by stage by water-supply"
```{r}

# Flag simulations that gave best fit (per crop only)
temp_data_bf <- statDF_crop_stage_irrig %>%
  mutate(BestFit = ifelse(index_crop %in% bestFit_dry$index_crop,"Best", "NotBest")) %>%
  mutate(BestFit = factor(BestFit, levels=c("NotBest", "Best")))  %>%
  filter(BestFit=="Best")

#summary(temp_data_bf)


```

Flag "behavioral" simulations (i.e. subjective "similarly-good" simulations) - NOT USED yet

```{r}

# FIXME: untested

# create df with "best fit" stat values and ranges
rangeBestFit <- 0.025 # % of value around rRMSE and LC

bound_bf <- temp_data_bf %>%
  filter(BestFit=="Best") %>%
  group_by(Crop, Stage) %>%
  dplyr::select(Crop, Stage,lc, r_rmse) %>%
  mutate(r_rmse_best = r_rmse, lc_best = lc) %>%
  dplyr::select(-lc, -r_rmse) %>%
  mutate(r_rmse_bestFit_low = (1-rangeBestFit)*r_rmse_best,
         r_rmse_bestFit_up = (1+rangeBestFit)*r_rmse_best,
         lc_bestFit_low = (1-rangeBestFit)*lc_best,
         lc_bestFit_up = (1+rangeBestFit)*lc_best)

summary(bound_bf)

```

```{r}
# Create a flag in big dataset to point if r_rmse and lc are within range of good fit
tempdf <- merge(statDF_crop_stage_irrig,bound_bf, by=c("Crop", "Stage"))

# quick in between function
is.between <- function(x, a, b) {
(x - a)  *  (b - x) > 0
}

# create a full dataset with flags for best fit boundaries
data_bf <- tempdf %>%
  rowwise() %>%
  mutate(r_rmse_range_flag = ifelse(is.between(r_rmse,r_rmse_bestFit_low,r_rmse_bestFit_up), 1, 0)) %>%
  mutate(lc_range_flag = ifelse(is.between(lc,lc_bestFit_low,lc_bestFit_up), 1, 0)) %>%
  mutate(range_idx = r_rmse_range_flag * lc_range_flag) %>%
  dplyr::select(-(r_rmse_range_flag:lc_range_flag),-(r_rmse_bestFit_low:lc_bestFit_up))
```

# Pin-point best-fit parameter within populations of simulations

- Note that the "best-fit" for dry is not necessarily the best fit for all 

```{r PlotSimPop, fig.height=20,fig.width=15, echo=FALSE, warning=FALSE}

data_bf %>%
  ggplot(aes(y=r_rmse, x=lc)) + # not working yet
  geom_point(alpha=0.8, aes(size=factor(KLReduction),
                 colour=factor(SurfaceKl),
                 shape=factor(RootFrontVelocity))) +
  scale_shape_discrete(solid=F)  +
  xlab("Lack of correlation (LC, % MSE) ") + ylab(" Relative RMSE (% observed mean)") +
  # labs(shape="Root Front Velocity (RFV)", colour="Surface kl (kls)", size="kl Reduction (klr)")+
  labs(shape="RFV", 
       colour=expression("kl"["s"]), 
       size=expression(lambda["kl"])) +
  theme(text = element_text(size=16))+
  theme(legend.position="top")  +
  facet_wrap(Crop+Stage~Irrigation,ncol=4) + # FIXME: Order matters - facet_grid might do the job?
  geom_vline(data=temp_data_bf,aes(xintercept=temp_data_bf$lc), linetype=2, colour="darkgrey") + 
  geom_hline(data=temp_data_bf, aes(yintercept=temp_data_bf$r_rmse), linetype=2, colour="darkgrey") 
  # now add elipse around "good-enough" simulations - FIXME 
 # + stat_ellipse(type = "norm", aes(linetype = range_idx == 1)) + 
 #  scale_linetype_manual(values=c("blank", "dotted")) 

```



