---
title: "MetFileAnalysis"
author: "edmar.teixeira@me.com"
date: "Friday, March 27, 2015"
output: html_document
---

## Reads met files from apsim (based on MapOutApsimFiles.R)
## Creates a dataframe with column row

```{r Libs, echo = FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(scales)

saveDir <- "\\\\Lin-file\\home$\\cfleit\\My Documents\\My papers\\2013-LucernePasture\\kl\\submitted\\round_6"
```

## The weather file

> That's how the data looks like
> EP calculated as Penman-Monteith by Hamish's Python scripts

```{r DirSet, echo=FALSE}

#Set paths
#setwd("C:\\Apsim_dev\\Projects\\Forage_MBIE(FRNL)\\ModelStudies")
#metFolder <- "C:\\Apsim_dev\\Data\\MetFiles\\"

metFolder <- "K:/CPDiary/Data/Lucerne & pasture drought trial 2011-2012/ModelFitting/"

# Get file
# metFile <- read.table(paste0(metFolder,"LincolnMetWithEP.txt"), 
#                       skip = 6, header = TRUE, comment.char = "(", blank.lines.skip = TRUE)

metFile <- read.table(paste0(metFolder,"LincolnMetWithEP.txt"), header = TRUE, sep=",") # revised by Hamish plus ep calculation

metFile <- data.frame(metFile)

summary(metFile)

# create a date
#metFile$Date <- as.Date(paste(metFile$year, metFile$day , sep = "." )  , format = "%Y.%j" )
metFile$Date <- as.Date(paste(metFile$year, metFile$day , sep = "." )  , format = "%Y.%j" )
metFile$month <- as.numeric(month(metFile$Date))


metFile %>%
  filter(ep < 0) %>%
  summary()





```

```{r}
metFile <- metFile %>%
  rowwise() %>%
  mutate(ep = ifelse(ep<0,0,as.numeric(as.character(ep))))
summary(metFile)
```

# Graph historical averages

> Daily temperaure and radiation

> monthly rainfall

> Pooled for 30 years

```{r GraphMet, echo = FALSE, fig.width=12, fig.height=10}

baseDate <- ymd("2011-01-15 UTC")

fontSize <- 15

# historical data df --------------------------------
weather_hist <- metFile %>%
  filter(year %in% 1960:2010) %>%
  group_by(month) %>%
  summarise_all(funs(mean, sd)) %>%
  as.data.frame()

# place holder for 4 similar sets of annual historical averages
df_exp_temp <- weather_hist[rep(row.names(weather_hist), 4), 1:ncol(weather_hist)]
df_exp_temp$month <- c(1:nrow(df_exp_temp)) # number the months sequentially

# create mid-point "reference" for trial period
weather_hist.expanded <- df_exp_temp %>%
  mutate(BaseDate = ymd(baseDate)) %>%
  mutate(Date =  BaseDate %m+% months(month-1)) %>% # creates dates starting from base and adding the number of months 
  dplyr::select(-BaseDate,-year_mean)

summary(weather_hist.expanded)

```
# test
```{r}
weather_hist.expanded %>%
  ggplot(aes(x=Date, y=maxt_mean)) +
  geom_line()
```

# do a daily historical average
```{r}
weather_day_avg <- metFile %>%
  filter(year %in% 1960:2010) %>%
#  mutate(DOY = yday(Date)) %>% # days of year
  group_by(day) %>%
  dplyr::select(day:month) %>%
  summarise_all(funs(mean)) %>%
  mutate(month = as.integer(month))%>%
  as.data.frame()

summary(weather_day_avg)
```
## create long term mean df
```{r}

str(df_exp_temp)

# Quick way of creating 3 similar years
y1 <- weather_day_avg %>% mutate(Date = ymd("2011-01-01 UTC") %m+% days(as.integer(day-1)))
y2 <- weather_day_avg %>% mutate(Date = ymd("2012-01-01 UTC") %m+% days(as.integer(day-1))) 
y3 <- weather_day_avg %>% mutate(Date = ymd("2013-01-01 UTC") %m+% days(as.integer(day-1)))
y4 <- weather_day_avg %>% mutate(Date = ymd("2014-01-01 UTC") %m+% days(as.integer(day-1)))

weather_day.hist <- NULL
weather_day.hist <- rbind(y1,y2,y3,y4)

# # FIXME: NOT working yet 
# 
# # place holder for multiplied historical averages
# #df_exp_temp <- weather_day_avg[rep(row.names(weather_day_avg), 4), 1:ncol(weather_day_avg)]
# #df_exp_temp$rowNo <- row.names.data.frame(df_exp_temp) 
# 
# # create mid-point "mock dates" for trial period
# weather_day.expanded <- df_exp_temp %>%
#   mutate(BaseDate = ymd(baseDate)) %>%
#   mutate(Date = ymd("2011-01-01 UTC") %m+% days(as.integer(day-1)))
#  # dplyr::select(-BaseDate,-year_mean)


```

```{r}
summary(weather_day.hist)
```

# select period of trials
```{r}
weather_trial <- metFile %>%
  filter(year %in% 2011:2014) %>%
  as.data.frame()

summary(weather_trial)
```

Get potential ET from APSIM

- now Hamish calculates in base file - not needed

```{r PotEvap_graph}

# Get file (simulated eo by APSIM - Pot ET for modsim paper)
# theFile <- "C:\\Apsim_dev\\Projects\\Forage_MBIE(FRNL)\\ModelStudies\\05_CoverCropPaperSoils\\LightSoil EarlyCover ETPdaily.out"
# 
# etpFile <- read.table(theFile, skip = 3, header = TRUE, comment.char = "(", blank.lines.skip = TRUE)
# # head(etpFile)
# # tail(etpFile)
# # summary(etpFile)
# 
# etp_work <- etpFile %>%
#   dplyr::select(Date:month, eo) %>%
#   mutate(Date = dmy(Date))
# 
# summary(etp_work)



```

# do averages of et pot

```{r}

# etp_day_av <- etp_work %>%
#   filter(year %in% 1970:2010) %>%
# #  mutate(DOY = yday(Date)) %>% # days of year
#   group_by(day_of_year) %>%
# #  dplyr::select(day:month) %>%
#   summarise_all(funs(mean)) %>%
#   mutate(month = as.integer(month))%>%
#   as.data.frame()
# 
# summary(weather_day_avg)


```

Append y similar seasons of histogram etp

```{r}
# # Quick way of creating 3 similar years
# y1 <- etp_day_av %>% mutate(Date = ymd("2011-01-01 UTC") %m+% days(as.integer(day_of_year-1)))
# y2 <- etp_day_av %>% mutate(Date = ymd("2012-01-01 UTC") %m+% days(as.integer(day_of_year-1))) 
# y3 <- etp_day_av %>% mutate(Date = ymd("2013-01-01 UTC") %m+% days(as.integer(day_of_year-1)))
# y4 <- etp_day_av %>% mutate(Date = ymd("2014-01-01 UTC") %m+% days(as.integer(day_of_year-1)))
# 
# etp_day_av.hist <- NULL
# etp_day_av.hist <- rbind(y1,y2,y3,y4)
# 
# summary(etp_day_av.hist)

```

Construct the composed 4 panel figure

```{r,fig.width=12, fig.height=8, warning=FALSE}

# maxt
g1 <- weather_trial %>%
  ggplot(aes(x=Date, y=maxt))  +
  geom_point(colour= "darkblue", alpha=0.3) +
  geom_line(data=weather_day.hist, 
            aes(x=Date, y=maxt), colour="black") +
 # theme_bw() +
  xlab("Date") + 
  ylab("Daily maximum temperature (oC)") +
  ylab(expression("Daily maximum temperature ( "*degree*C*")")) +
  theme(text = element_text(size=fontSize)) 

#mint
g2 <- weather_trial %>%
  ggplot(aes(x=Date, y=mint))  +
  geom_point(colour= "darkblue", alpha=0.3) +
  geom_line(data=weather_day.hist, 
            aes(x=Date, y=mint), colour="black") +
 # theme_bw() +
  xlab("Date") + 
  ylab("Daily minimum temperature (oC)") +
  ylab(expression("Daily minimum temperature ( "*degree*C*")")) +
  theme(text = element_text(size=fontSize)) 

#radn
g3 <- weather_trial %>%
  ggplot(aes(x=Date, y=radn))  +
  geom_point(colour= "darkblue", alpha=0.3) +
  geom_line(data=weather_day.hist, 
            aes(x=Date, y=radn), colour="black") +
 # theme_bw() +
  xlab("Date") + 
  ylab("Daily maximum temperature (oC)") +
#  ylab(expression("Daily maximum temperature ( "*degree*C*")")) +
 # ylab("Daily solar radiation (MJ/m2)" ) +
   ylab(expression(Daily~solar~radiation~~"(MJ "*m^"-2"*")")) +
   theme(text = element_text(size=fontSize))

# rain 
#weather_trial$month <- with(weather_trial, strptime(paste(year, day), format="%Y %j")$mon+1)
rain.month.hyst <- weather_day.hist %>% 
  mutate(year=year(Date)) %>%
  group_by(month, year)  %>%
  summarise(rain= sum(rain) , date = mean(Date), ep = sum(ep))

g4 <- weather_trial %>%
  group_by(month,year) %>%
  summarise(rain= sum(rain) , date = mean(Date), ep=sum(ep)) %>%
  # ungroup() +
  # group_by(month,year) %>%
  # select(month,year, rain) %>%
#  summarise(rainMean = mean(rain), rainSd = sd(rain)/2) %>%
  ggplot(aes(x=date, y=rain)) + 
  geom_bar(stat ='identity', fill="white", colour="black") +  
  # geom_line(data=rain.month.hyst, 
  #           aes(x=date, y=rain), colour="black") +
    # geom_line(data=weather_day.hist, 
    #         aes(x=Date, y=rain*30.5), colour="blue", alpha=0.3) +
      # geom_line(data=rain.month.hyst, 
   geom_line(data=rain.month.hyst, aes(x=date, y=ep), colour="grey", linetype=2) +
   geom_line(data=rain.month.hyst, aes(x=date, y=rain), colour="black", linetype=2) +
   geom_point(aes(x=date, y=ep), colour= "darkblue", alpha=0.3) +
 # theme_bw() +
  xlab("Date") + 
  ylab("Monthly rainfall and ET (mm)") +
 # scale_x_continuous(breaks=1:12) +
  # geom_errorbar(aes(ymax = rainMean + rainSd, ymin=rainMean - rainSd), 
  #               position=position_dodge(width=0.9), width=0.25)+
  theme(text = element_text(size=fontSize))


g <- arrangeGrob(g1, g2, g3, g4, nrow=2)
g


# Save
 ggsave(file="Fig_Weather.tiff", dpi = 300, path=saveDir, g)

```


```{r}
# g5 <- etpFile %>%
#   filter(year %in% 1960:2000) %>%
#    group_by(day_of_year) %>%
#  # summarise_all(funs(mean)) %>%
#   ggplot(aes(x=day_of_year, y=eo))  +
#   geom_point(colour= "darkblue", alpha=0.2) +
#   stat_smooth(level = 0.9, size = 1, colour="white") + 
#   xlab("Day of the year (1-365)") + 
#   ylab("Potential evapo-transpiration (mm)") +
#   theme(text = element_text(size=fontSize))
# 
# 
# g5 <- etpFile %>%
#   filter(year %in% 1960:2000) %>%
#   group_by(year, month) %>%
#   summarise_all(funs(sum)) %>%
#   group_by(month) %>%
#   select(month, eo) %>%
#   summarise(rainMean = mean(eo), rainSd = sd(eo)/2) %>%
#   ggplot(aes(x=month, y=rainMean)) + 
#   geom_bar(stat ='identity') +
#   xlab("month") + 
#   ylab("Potential evapo-transpiration (mm)") +
#   scale_x_continuous(breaks=1:12) +
#   geom_errorbar(aes(ymax = rainMean + rainSd, ymin=rainMean - rainSd), 
#                 position=position_dodge(width=0.9), width=0.25)+
#   theme(text = element_text(size=fontSize))
# 
# 
# grid.arrange(g1, g3, g4, g5)

```


